### 2018-05-31 article_code.Rmd
added the code

---
title: "R Notebook"
output: html_notebook
---

# the project is in Dropbox/coding/cosimulation

```{r}
source("C:/Users/tl3315/Dropbox/coding/cosimulation/co_sim_condi.R")
source("C:/Users/tl3315/Dropbox/coding/cosimulation/cosim_functions.R")
```
```{r}
library(mvtnorm)
library(abind)
```

```{r}
e1<-e2<-e3<-e4<-d1<-d2<-d3<-matrix(0,nz,4)
e1[env_lat==1]=1
e2[env_lat==2]=1
e3[env_lat==3]=1
e4[env_lat==4]=1
d1[diag_lat==1]=1
d2[diag_lat==2]=1
d3[diag_lat==3]=1
e1[is.na(env_lat)]=e2[is.na(env_lat)]=e3[is.na(env_lat)]=e4[is.na(env_lat)]=NA
d1[is.na(diag_lat)]=d2[is.na(diag_lat)]=d3[is.na(diag_lat)]=NA
```
```{r}
pe1 = mean(e1,na.rm=TRUE)
pe2 = mean(e2,na.rm=TRUE)
pe3 = mean(e3,na.rm=TRUE)
pe4 = mean(e4,na.rm=TRUE)
c(pe1,pe2,pe3,pe4)
pd1=mean(d1,na.rm=TRUE)
pd2=mean(d2,na.rm=TRUE)
pd3=mean(d3,na.rm=TRUE)
c(pd1,pd2,pd3)
pd1e1=mean(d1*e1,na.rm=TRUE)/pe1
pd1e2=mean(d1*e2,na.rm=TRUE)/pe2
pd1e3=mean(d1*e3,na.rm=TRUE)/pe3
pd1e4=mean(d1*e4,na.rm=TRUE)/pe4

pd2e1=mean(d2*e1,na.rm=TRUE)/pe1
pd2e2=mean(d2*e2,na.rm=TRUE)/pe2
pd2e3=mean(d2*e3,na.rm=TRUE)/pe3
pd2e4=mean(d2*e4,na.rm=TRUE)/pe4

pd3e1=mean(d3*e1,na.rm=TRUE)/pe1
pd3e2=mean(d3*e2,na.rm=TRUE)/pe2
pd3e3=mean(d3*e3,na.rm=TRUE)/pe3
pd3e4=mean(d3*e4,na.rm=TRUE)/pe4

matrix(c(pd1e1,pd1e2,pd1e3,pd1e4,pd2e1,pd2e2,pd2e3,pd2e4,pd3e1,pd3e2,pd3e3,pd3e4),3,4,byrow=TRUE)
pe1+pe2+pe3+pe4
```

```{r}
env_lat[is.na(env_lat)]=0 # with 0, the gibbs sampler will find random values
diag_lat[is.na(diag_lat)]=0
grid<-grid_co<-array(NA,c(100,10,91))
grid[5,8,]<-env_lat[,1]
grid[25,6,]<-env_lat[,2]
grid[35,3,]<-env_lat[,3]
grid[95,7,]<-env_lat[,4]
grid_co[5,8,]<-diag_lat[,1]
grid_co[25,6,]<-diag_lat[,2]
grid_co[35,3,]<-diag_lat[,3]
grid_co[95,7,]<-diag_lat[,4]

```
ok we also want to display an image with all the logs
we are doubling the wells for a better visualization
```{r}
im<-matrix(NA,100,91)
im[5,]<-env_lat[,1]
im[4,]<-env_lat[,1]
im[25,]<-env_lat[,2]
im[24,]<-env_lat[,2]
im[35,]<-env_lat[,3]
im[34,]<-env_lat[,3]
im[95,]<-env_lat[,4]
im[94,]<-env_lat[,4]
image(x=(1:100)*9.2,y=(1:91)/10,z=im,col=c("lightgreen","dodgerblue4","darkmagenta","red"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```
print image
```{r}
png(file="logslate.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*9.2,y=(1:91)/10,z=im,col=c("lightgreen","dodgerblue4","darkmagenta","red"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```
we also want to display the environment of depositions and the diagenesis
```{r}
im<-matrix(NA,100,91)
im[5,]<-diag_lat[,1]+4
im[4,]<-diag_lat[,1]+4
im[8,]<-env_lat[,1]
im[7,]<-env_lat[,1]
im[25,]<-diag_lat[,2]+4
im[24,]<-diag_lat[,2]+4
im[28,]<-env_lat[,2]
im[27,]<-env_lat[,2]
im[35,]<-diag_lat[,3]+4
im[34,]<-diag_lat[,3]+4
im[38,]<-env_lat[,3]
im[37,]<-env_lat[,3]
im[95,]<-diag_lat[,4]+4
im[94,]<-diag_lat[,4]+4
im[98,]<-env_lat[,4]
im[97,]<-env_lat[,4]
image(x=(1:100)*9.2,y=(1:91)/10,z=im,col=c("lightgreen","dodgerblue4","darkmagenta","red","black","grey","yellow"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```
```{r}
png(file="logslatediag2.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*9.2,y=(1:91)/10,z=im,col=c("lightgreen","dodgerblue4","darkmagenta","red","black","grey","yellow"),ylab="Vertical distance (m)",xlab = "Lateral distance (m)")
```

```{r}
grid2<-grid_co2<-array(NA,c(100,10,182))
i=1
for (j in 1:182){

    grid2[5,8,j]<-env_lat[i,1]
    grid2[25,6,j]<-env_lat[i,2]
    grid2[35,3,j]<-env_lat[i,3]
    grid2[95,7,j]<-env_lat[i,4]
    grid_co2[5,8,j]<-diag_lat[i,1]
    grid_co2[25,6,j]<-diag_lat[i,2]
    grid_co2[35,3,j]<-diag_lat[i,3]
    grid_co2[95,7,j]<-diag_lat[i,4]
    if(j%%2==0){i=i+1}
}
```

the first figure we want to have is a 3d simulation of the Latemar conditioned ! 
```{r}
rv1=0.3
rv2=1.2
rhx1=rhy1=800
rhx2=rhy2=800
b1=0
b2=5
rho=0.6
shift=0.1
rho1a<-model(shift,rv1,b1)
warning(rho>rho1a)
alpha1=(rho/rho1a)^2
alpha2=1-(rho/rho1a)^2
rv3=0.2
rv4=0.3
rhx3=rhy3=800
rhx4=rhy4=800
b3=0
b4=0
e=-0.8
f=-0.5
a14=-0.1
a24=0.1
rho1a14 = model(a14,rv1,b1)
rho2a24=model(a24,rv2,b2)

#thresholds
q1=qnorm(pe1+pe2)
Sigma1 = matrix(c(1,rho,rho,1),2,2)
q2=0
obj = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1,q2test),mean=c(0,0),sigma=Sigma1) - pe1 ) 
  if (newobj < obj){
    q2 = q2test
    obj = newobj
  }
}
q3=0
obj = Inf
for ( test in 1:100 ){
  q3test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q3test),mean=c(0,0),sigma=Sigma1) - pe4 ) 
  if (newobj < obj){
    q3 = q3test
    obj = newobj
  }
}
c(q1,q2,q3)
qd1e4=qnorm(pd3e4) # no correlation with the third Gaussian function
Sigmad = diag(1,3)
qd2e2=0
qd2e3=0
obje2 = Inf
obje3 = Inf
for ( test in 1:100 ){
  qd2e2test = qnorm(test/100)
  qd2e3test = qnorm(test/100)
  Sigmad[1,2]=Sigmad[2,1]=e*rho1a14
  Sigmad[1,3]=Sigmad[3,1]=e*(rho/rho1a)*model(a14-shift,rv1,b1)+f*sqrt(1-(rho/rho1a)^2)*model(a24,rv2,b2)
  Sigmad[2,3]=Sigmad[3,2]=rho
  newobje2 =abs( pmvnorm(lower=c(qd2e2test,-Inf,q2),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=Sigmad)/pe2 - pd2e2 ) 
  newobje3 =abs( pmvnorm(lower=c(qd2e3test,q1,q3),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=Sigmad)/pe3 - pd2e3 )
  if (newobje2 < obje2){
    qd2e2 = qd2e2test
    obje2 = newobje2
  }
  if (newobje3 < obje3){
    qd2e3 = qd2e3test
    obje3 = newobje3
  }
}
c(qd1e4,qd2e2,qd2e3)

Y1<-sim3d(rv1,rhx1,rhy1,b1,100,0.1,91,10,100,25,10)
Y2<-sim3d(rv2,rhx2,rhy2,b2,100,0.1,91,10,100,25,10)
Y3<-sim3d(rv3,rhx3,rhy3,b3,100,0.1,91,10,100,25,10)
Y4<-sim3d(rv4,rhx4,rhy4,b4,100,0.1,91,10,100,25,10)

simwell<-gibbs_sampling3d(grid,grid_co,c(q1,q2,q3),c(qd1e4,qd2e2,qd2e3),0.1,10,25,c(rv1,rv2,rv3,rv4),c(rhx1,rhx2,rhx3,rhx4),c(rhy1,rhy2,rhy3,rhy4),c(b1,b2,b3,b4),c(rho,e*model(a14,rv1,b1),f*model(a24,rv2,b2)),c(shift,a14,a24),50)
Y1well<-simwell[,,1:91]
Y2well<-Z2well<-simwell[,,92:182]
Y3well<-simwell[,,183:273]
Y4well<-Z4well<-simwell[,,274:364]

# back transformation of the Zi into Yi at the wells
for( z in 2:90){
  Y2well[,,z]=(Z2well[,,z]-(rho/rho1a)*Y1well[,,z+1])/sqrt(1-(rho^2)/(rho1a^2))
  Y4well[,,z]=(Z4well[,,z]-Y1well[,,z-1]*e-f*Y2well[,,z+1])/sqrt(1-(e^2)-(f^2))
}

Z1c =Y1 + dual_krig_surf(Y1well-Y1,rhx1,rhy1,10,25)
Z2c<-Y2c<-Y2 +dual_krig_surf(Y2well-Y2,rhx2,rhy2,10,25)
Z3c<-Y3 +dual_krig_surf(Y3well-Y3,rhx3,rhy3,10,25)
Z4c<-Y4c<-Y4 +dual_krig_surf(Y4well-Y4,rhx4,rhy4,10,25)

for( z in 2:90){
  Z2c[,,z]=Y2c[,,z]*sqrt(1-(rho^2)/(rho1a^2))+Z1c[,,z+1]*(rho)/(rho1a)
  Z4c[,,z]=Y4c[,,z]*sqrt(1-(e^2)-(f^2))+Z1c[,,z-1]*e+f*Y2c[,,z+1]
}

faciesenv<-faciesdiag<-Z1c
faciesenv[Z1c<q1 & Z2c<q2]=1
faciesenv[Z1c<q1 & Z2c>q2]=2
faciesenv[Z1c>q1 & Z2c<q3]=4
faciesenv[Z1c>q1 & Z2c>q3]=3
faciesdiag[]=1
faciesdiag[Z1c>q1 & Z2c<q3 & Z3c<qd1e4]=3
faciesdiag[Z1c<q1 & Z2c>q2 & Z4c>qd2e2]=2
faciesdiag[Z1c>q1 & Z2c>q3 & Z4c>qd2e3]=2
col_env <- c("lightgreen","cadetblue","dodgerblue4","lightcoral")
griddata<-grid
griddata[griddata==0]=NA
W<-image(griddata[,2,],col = col_env)
W<-image(faciesenv[,2,],col = col_env)

```

```{r}
for ( z in 1:91){
  for ( y in 1:10){
    for ( x in 1:100){
      write(toString(c(faciesenv[x,y,z])),"latemarcondenv.csv",append=TRUE)
      write(toString(c(faciesdiag[x,y,z])),"latemarconddiag.csv",append=TRUE)
    }
  }
}
```
another simu with only three Gaussian functions
(verifier le gibbs sampling pour la nouvelle truncation rule)

```{r}
library(mvtnorm)
```
```{r}
source("co_sim_condi.R")
```

```{r}
par(bg="transparent")
rv1=0.3
rv2=1.2
rhx1=rhy1=800
rhx2=rhy2=800
b1=0
b2=5
rho=0.6
shift=0.1
rho1a<-model(shift,rv1,b1)
warning(rho>rho1a)
alpha1=(rho/rho1a)^2
alpha2=1-(rho/rho1a)^2
rv3=0.2
rv4=0.3
rhx3=rhy3=800
rhx4=rhy4=800
b3=0
b4=0
e=-0.8
f=-0.5
a14=-0.1
a24=0.1
rho1a14 = model(a14,rv1,b1)
rho2a24=model(a24,rv2,b2)

#thresholds
q1=qnorm(pe1+pe2)
Sigma1 = matrix(c(1,rho,rho,1),2,2)
q2=0
obj = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1,q2test),mean=c(0,0),sigma=Sigma1) - pe1 ) 
  if (newobj < obj){
    q2 = q2test
    obj = newobj
  }
}
q3=0
obj = Inf
for ( test in 1:100 ){
  q3test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q3test),mean=c(0,0),sigma=Sigma1) - pe4 ) 
  if (newobj < obj){
    q3 = q3test
    obj = newobj
  }
}
c(q1,q2,q3)

Sigmad = diag(1,3)
qd2e2=0
qd2e3=0
qd1e4=0
obje2 = Inf
obje3 = Inf
obje4 = Inf
for ( test in 1:100 ){
  qd2e2test = qnorm(test/100)
  qd2e3test = qnorm(test/100)
  qd1e4test = qnorm(test/100)
  Sigmad[1,2]=Sigmad[2,1]=e*rho1a14
  Sigmad[1,3]=Sigmad[3,1]=e*(rho/rho1a)*model(a14-shift,rv1,b1)+f*sqrt(1-(rho/rho1a)^2)*model(a24,rv2,b2)
  Sigmad[2,3]=Sigmad[3,2]=rho
  newobje2 =abs( pmvnorm(lower=c(qd2e2test,-Inf,q2),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=Sigmad)/pe2 - pd2e2 ) 
  newobje3 =abs( pmvnorm(lower=c(qd2e3test,q1,q3),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=Sigmad)/pe3 - pd2e3 )
  newobje4 =abs( pmvnorm(lower=c(qd1e4test,q1,-Inf),upper=c(Inf,Inf,q3),mean=c(0,0,0),sigma=Sigmad)/pe4 - pd3e4 )
  if (newobje2 < obje2){
    qd2e2 = qd2e2test
    obje2 = newobje2
  }
  if (newobje3 < obje3){
    qd2e3 = qd2e3test
    obje3 = newobje3
  }
  if (newobje4 < obje4){
    qd1e4 = qd1e4test
    obje4 = newobje4
  }
}

c(qd1e4,qd2e2,qd2e3)
# here Y4 is not used
Y1<-sim3d(rv1,rhx1,rhy1,b1,100,0.05,182,10,100,25,10)
Y2<-sim3d(rv2,rhx2,rhy2,b2,100,0.05,182,10,100,25,10)
Y3<-sim3d(rv3,rhx3,rhy3,b3,100,0.05,182,10,100,25,10)
Y4<-sim3d(rv4,rhx4,rhy4,b4,100,0.05,182,10,100,25,10)
#careful to check that we put 0 to extend the gibbs sampling to where there is no data
simwell<-gibbs_sampling3d(grid2,grid_co2,c(q1,q2,q3),c(qd1e4,qd2e2,qd2e3),0.05,10,25,c(rv1,rv2,rv3,rv4),c(rhx1,rhx2,rhx3,rhx4),c(rhy1,rhy2,rhy3,rhy4),c(b1,b2,b3,b4),c(rho,e*model(a14,rv1,b1),f*model(a24,rv2,b2)),c(shift,a14,a24),50)
Y1well<-simwell[,,1:182]
Y2well<-Z2well<-simwell[,,183:364]
Y3well<-simwell[,,365:546]
Y4well<-Z4well<-simwell[,,547:728]

# back transformation of the Zi into Yi at the wells
# be VERY CAREFUL, we cannot put Y4 and Y2 in the same loop
for( z in 3:180){
  Y2well[,,z]=(Z2well[,,z]-(rho/rho1a)*Y1well[,,z+2])/sqrt(1-(rho^2)/(rho1a^2))
}
for( z in 3:180){
  Y4well[,,z]=(Z4well[,,z]-Y1well[,,z-2]*e-f*Y2well[,,z+2])/sqrt(1-(e^2)-(f^2))
}

Z1c =Y1 + dual_krig_surf(Y1well-Y1,rhx1,rhy1,10,25)
Z2c<-Y2c<-Y2 +dual_krig_surf(Y2well-Y2,rhx2,rhy2,10,25)
Z3c<-Y3 +dual_krig_surf(Y3well-Y3,rhx3,rhy3,10,25)
Z4c<-Y4c<-Y4 +dual_krig_surf(Y4well-Y4,rhx4,rhy4,10,25)

for( z in 3:180){
  Z2c[,,z]=Y2c[,,z]*sqrt(1-(rho^2)/(rho1a^2))+Z1c[,,z+2]*(rho)/(rho1a)
}
for( z in 3:180){
  Z4c[,,z]=Y4c[,,z]*sqrt(1-(e^2)-(f^2))+Z1c[,,z-2]*e+f*Y2c[,,z+2]
}

faciesenv<-faciesdiag<-Z1c
faciesenv[Z1c<q1 & Z2c<q2]=1
faciesenv[Z1c<q1 & Z2c>q2]=2
faciesenv[Z1c>q1 & Z2c<q3]=4
faciesenv[Z1c>q1 & Z2c>q3]=3
faciesdiag[]=1
faciesdiag[Z1c>q1 & Z2c<q3 & Z4c>qd1e4]=3
faciesdiag[Z1c<q1 & Z2c>q2 & Z4c>qd2e2]=2
faciesdiag[Z1c>q1 & Z2c>q3 & Z4c>qd2e3]=2
col_env <- c("lightgreen","cadetblue","dodgerblue4","lightcoral")
griddata<-grid2
griddata[griddata==0]=NA
W<-image(griddata[,8,],col = col_env)
W<-image(faciesenv[,8,],col = col_env)
W<-image(grid_co2[,8,])
W<-image(faciesdiag[,8,])
```
computing transiograms on the realization
```{r}
i1real<-i2real<-i3real<-i4real<-faciesenv
i1real[faciesenv!=1]=0
i2real[faciesenv==2]=1
i2real[faciesenv!=2]=0
i3real[faciesenv==3]=1
i3real[faciesenv!=3]=0
i4real[faciesenv==4]=1
i4real[faciesenv!=4]=0
t11real3=tr_3d_vert(i1real,i1real,60)
t12real3=tr_3d_vert(i1real,i2real,60)
t13real3=tr_3d_vert(i1real,i3real,60)
t14real3=tr_3d_vert(i1real,i4real,60)
t21real3=tr_3d_vert(i2real,i1real,60)
t22real3=tr_3d_vert(i2real,i2real,60)
t23real3=tr_3d_vert(i2real,i3real,60)
t24real3=tr_3d_vert(i2real,i4real,60)
t31real3=tr_3d_vert(i3real,i1real,60)
t32real3=tr_3d_vert(i3real,i2real,60)
t33real3=tr_3d_vert(i3real,i3real,60)
t34real3=tr_3d_vert(i3real,i4real,60)
t41real3=tr_3d_vert(i4real,i1real,60)
t42real3=tr_3d_vert(i4real,i2real,60)
t43real3=tr_3d_vert(i4real,i3real,60)
t44real3=tr_3d_vert(i4real,i4real,60)

z = (1:60)/20.
png(file="laterealtransio.png",width=2000,height = 1600,res=300)
par(mar=c(2,3,1,1),mfrow=c(4,4))
plot(z,t11real)
title(main=expression(paste('t'[e1e1],"(h)")),line=-1)
plot(z,t12real)
title(main=expression(paste('t'[e1e2],"(h)")),line=-1)
plot(z,t13real)
title(main=expression(paste('t'[e1e3],"(h)")),line=-1)
plot(z,t14real)
title(main=expression(paste('t'[e1e4],"(h)")),line=-1)
plot(z,t21real)
title(main=expression(paste('t'[e2e1],"(h)")),line=-1)
plot(z,t22real)
title(main=expression(paste('t'[e2e2],"(h)")),line=-1)
plot(z,t23real)
title(main=expression(paste('t'[e2e3],"(h)")),line=-1)
plot(z,t24real)
title(main=expression(paste('t'[e2e4],"(h)")),line=-1)
plot(z,t31real)
title(main=expression(paste('t'[e3e1],"(h)")),line=-1)
plot(z,t32real)
title(main=expression(paste('t'[e3e2],"(h)")),line=-1)
plot(z,t33real)
title(main=expression(paste('t'[e3e3],"(h)")),line=-1)
plot(z,t34real)
title(main=expression(paste('t'[e3e4],"(h)")),line=-1)
plot(z,t41real)
title(main=expression(paste('t'[e4e1],"(h)")),line=-1)
plot(z,t42real)
title(main=expression(paste('t'[e4e2],"(h)")),line=-1)
plot(z,t43real)
title(main=expression(paste('t'[e4e3],"(h)")),line=-1)
plot(z,t44real)
title(main=expression(paste('t'[e4e4],"(h)")),line=-1)
```
average
```{r}
t11reals=(t11real+t11real2+t11real3)/3
t12reals=(t12real+t12real2+t12real3)/3
t13reals=(t13real+t13real2+t13real3)/3
t14reals=(t14real+t14real2+t14real3)/3
t21reals=(t21real+t21real2+t21real3)/3
t22reals=(t22real+t22real2+t22real3)/3
t23reals=(t23real+t23real2+t23real3)/3
t24reals=(t24real+t24real2+t24real3)/3
t31reals=(t31real+t31real2+t31real3)/3
t32reals=(t32real+t32real2+t32real3)/3
t33reals=(t33real+t33real2+t33real3)/3
t34reals=(t34real+t34real2+t34real3)/3
t41reals=(t41real+t41real2+t41real3)/3
t42reals=(t42real+t42real2+t42real3)/3
t43reals=(t43real+t43real2+t43real3)/3
t44reals=(t44real+t44real2+t44real3)/3
```
and also with the diagenetic facies
```{r}
d1real<-d2real<-d3real<-faciesdiag
d1real[faciesdiag!=1]=0
d2real[faciesdiag==2]=1
d2real[faciesdiag!=2]=0
d3real[faciesdiag==3]=1
d3real[faciesdiag!=3]=0
td3e1real3=tr_3d_vert(d3real,i1real,60)
td3e2real3=tr_3d_vert(d3real,i2real,60)
td3e3real3=tr_3d_vert(d3real,i3real,60)
td3e4real3=tr_3d_vert(d3real,i4real,60)
td2e1real3=tr_3d_vert(d2real,i1real,60)
td2e2real3=tr_3d_vert(d2real,i2real,60)
td2e3real3=tr_3d_vert(d2real,i3real,60)
td2e4real3=tr_3d_vert(d2real,i4real,60)
td3d3real3=tr_3d_vert(d3real,d3real,60)
td2d3real3=tr_3d_vert(d2real,d3real,60)
td3d2real3=tr_3d_vert(d3real,d2real,60)
td2d2real3=tr_3d_vert(d2real,d2real,60)

```

```{r}
for ( z in 1:182){
  for ( y in 1:10){
    for ( x in 1:100){
      write(toString(c(faciesenv[x,y,z])),"latemarcondenvres6.csv",append=TRUE)
      write(toString(c(faciesdiag[x,y,z])),"latemarconddiagres.csv",append=TRUE)
    }
  }
}
```
we can also combine the environment and the diagenesis in a single model
```{r}
combined<-faciesenv
combined[faciesdiag==2]=5
combined[faciesdiag==3]=6
for ( z in 1:182){
  for ( y in 1:10){
    for ( x in 1:100){
      write(toString(c(combined[x,y,z])),"latemarcombined4.csv",append=TRUE)
    }
  }
}
```

we also want to enter the data ! 
```{r}
griddata<-grid
griddata[griddata==0]=NA
gridcodata<-grid_co
gridcodata[gridcodata==0]=NA

for ( z in 1:91){
  for ( y in 1:10){
    for ( x in 1:100){
      write(toString(c(griddata[x,y,z])),"dataenv.csv",append=TRUE)
      write(toString(c(gridcodata[x,y,z])),"datadiag.csv",append=TRUE)
    }
  }
}
```
data with all facies
```{r}
griddata<-grid
griddata[griddata==0]=NA
gridcodata<-grid_co
gridcodata[gridcodata==0]=NA
combineddata<-griddata
combineddata[gridcodata==2]=5
combineddata[gridcodata==3]=6
for ( z in 1:91){
  for ( y in 1:10){
    for ( x in 1:100){
      write(toString(c(combineddata[x,y,z])),"latemarcombineddata.csv",append=TRUE)
    }
  }
}
```
let s compute the embedded transitions of one realization
```{r}
i1real<-i2real<-i3real<-i4real<-faciesenv
i1real[faciesenv!=1]=0
i2real[faciesenv==2]=1
i2real[faciesenv!=2]=0
i3real[faciesenv==3]=1
i3real[faciesenv!=3]=0
i4real[faciesenv==4]=1
i4real[faciesenv!=4]=0
r12=embedded3d(i1real,i2real)
r13=embedded3d(i1real,i3real)
r14=embedded3d(i1real,i4real)
r21=embedded3d(i2real,i1real)
r23=embedded3d(i2real,i3real)
r24=embedded3d(i2real,i4real)
r31=embedded3d(i3real,i1real)
r32=embedded3d(i3real,i2real)
r34=embedded3d(i3real,i4real)
r41=embedded3d(i4real,i1real)
r42=embedded3d(i4real,i2real)
r43=embedded3d(i4real,i3real)
matrix(c(0,r12,r13,r14,r21,0,r23,r24,r31,r32,0,r34,r41,r42,r43,0),4,4,byrow=TRUE)
```
```{r}
real7<-matrix(c(0,r12,r13,r14,r21,0,r23,r24,r31,r32,0,r34,r41,r42,r43,0),4,4,byrow=TRUE)
```
```{r}
(real1+real2+real3+real4+real5+real6)/6
```

and the data
```{r}
r12d=embedded2d(e1,e2)
r13d=embedded2d(e1,e3)
r14d=embedded2d(e1,e4)
s1=r12d+r13d+r14d
r21d=embedded2d(e2,e1)
r23d=embedded2d(e2,e3)
r24d=embedded2d(e2,e4)
s2=r21d+r23d+r24d
r31d=embedded2d(e3,e1)
r32d=embedded2d(e3,e2)
r34d=embedded2d(e3,e4)
s3=r31d+r32d+r34d
r41d=embedded2d(e4,e1)
r42d=embedded2d(e4,e2)
r43d=embedded2d(e4,e3)
s4=r41d+r42d+r43d
matrix(c(0,r12d/s1,r13d/s1,r14d/s1,r21d/s2,0,r23d/s2,r24d/s2,r31d/s3,r32d/s3,0,r34d/s3,r41d/s4,r42d/s4,r43d/s4,0),4,4,byrow=TRUE)
```

ok it is also necessary to compute the transiograms on the final model
```{r}
real1<-real2<-real3<-real4<-faciesenv
real1[faciesenv!=1]=0
real2[faciesenv==2]=1
real2[faciesenv!=2]=0
real3[faciesenv==3]=1
real3[faciesenv!=3]=0
real4[faciesenv==4]=1
real4[faciesenv!=4]=0
t11real<-tr_3d_vert(real1,real1,50)
t12real<-tr_3d_vert(real1,real2,50)
t13real<-tr_3d_vert(real1,real3,50)
t14real<-tr_3d_vert(real1,real4,50)
t21real<-tr_3d_vert(real2,real1,50)
t22real<-tr_3d_vert(real2,real2,50)
t23real<-tr_3d_vert(real2,real3,50)
t24real<-tr_3d_vert(real2,real4,50)
t31real<-tr_3d_vert(real3,real1,50)
t32real<-tr_3d_vert(real3,real2,50)
t33real<-tr_3d_vert(real3,real3,50)
t34real<-tr_3d_vert(real3,real4,50)
t41real<-tr_3d_vert(real4,real1,50)
t42real<-tr_3d_vert(real4,real2,50)
t43real<-tr_3d_vert(real4,real3,50)
t44real<-tr_3d_vert(real4,real4,50)
par(mfrow=c(4,4))
plot(50,t11real)
plot(length(t44real),t12real)
plot(length(t44real),t13real)
plot(length(t44real),t14real)
plot(length(t44real),t21real)
plot(length(t44real),t22real)
plot(length(t44real),t23real)
plot(length(t44real),t24real)
plot(length(t44real),t31real)
plot(length(t44real),t32real)
plot(length(t44real),t33real)
plot(length(t44real),t34real)
plot(length(t44real),t41real)
plot(length(t44real),t42real)
plot(length(t44real),t43real)
plot(length(t44real),t44real)
```

ok now that we have looked at the case study of the Latemar, 
we also want to create synthetic examples with a simpler parameter space. 
So we have only two Gaussian functions with two truncation rules, so it is like two TGS!

```{r}
model<-function(h,range,b){return(exp(-abs(h/range)^2)*cos(b*h) )}
```
```{r}
library(mvtnorm)
```

```{r}

# set of parameters 
pe1=0.5
pe2=1-pe1
pe1d1=0.0
pe2d1=0.3
rv1=0.3
rv2=1
rhx1=40
rhx2=10
b1=0
b2=0
a=0.9
shift=0.1
rho1a<-model(shift,rv1,b1)

#thresholds
q1=qnorm(pe1) # threshold for environment of deposition

q2e1=q2e2=0 # two thresholds for diagenetic facies 
Sigma = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
obj1=obj2=Inf
for ( test in 1:100 ){
  q2e1test = qnorm(test/100)
  q2e2test = qnorm(test/100)
  newobj1 =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1,q2e1test),mean=c(0,0),sigma=Sigma)/pe1 - pe1d1) 
  newobj2 =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q2e2test),mean=c(0,0),sigma=Sigma)/pe2 - pe2d1)
  if (newobj1 < obj1){
    q2e1= q2e1test
    obj1 = newobj1
  }
  if (newobj2 < obj2){
    q2e2= q2e2test
    obj2 = newobj2
  }
}

#transiograms
te1d1=te2d1=te1d2=te2d2=rep(0,300)
te1d1opp=te2d1opp=te1d2opp=te2d2opp=rep(0,300)
Sigma = diag(1,3)
Sigma[3,2]=Sigma[2,3]=a*rho1a
Sigmaopp = Sigma

for (i in 1:300){
    d = i/100.
    Sigma[1,2]=Sigma[2,1]= model(d,rv1,b1)
    Sigma[1,3]=Sigma[3,1]=a*model(d+shift,rv1,b1)

    Sigmaopp[1,2]=Sigmaopp[2,1]= model(-d,rv1,b1)
    Sigmaopp[1,3]=Sigmaopp[3,1]=a*model(-d+shift,rv1,b1)

    te1d1[i]=(pmvnorm(lower=c(-Inf,-Inf,-Inf),upper=c(q1,q1,q2e1),mean=c(0,0,0),sigma=Sigma) + pmvnorm(lower=c(-Inf,q1,-Inf),upper=c(q1,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe1
    te2d1[i]=(pmvnorm(lower=c(q1,-Inf,-Inf),upper=c(Inf,q1,q2e1),mean=c(0,0,0),sigma=Sigma) + pmvnorm(lower=c(q1,q1,-Inf),upper=c(Inf,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe2
    te1d2[i]=(pmvnorm(lower=c(-Inf,-Inf,q2e1),upper=c(q1,q1,Inf),mean=c(0,0,0),sigma=Sigma) + pmvnorm(lower=c(-Inf,q1,q2e2),upper=c(q1,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe1
    te2d2[i]=(pmvnorm(lower=c(q1,-Inf,q2e1),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=Sigma) + pmvnorm(lower=c(q1,q1,q2e2),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe2

    te1d1opp[i]=(pmvnorm(lower=c(-Inf,-Inf,-Inf),upper=c(q1,q1,q2e1),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(-Inf,q1,-Inf),upper=c(q1,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe1
    te2d1opp[i]=(pmvnorm(lower=c(q1,-Inf,-Inf),upper=c(Inf,q1,q2e1),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(q1,q1,-Inf),upper=c(Inf,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe2
    te1d2opp[i]=(pmvnorm(lower=c(-Inf,-Inf,q2e1),upper=c(q1,q1,Inf),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(-Inf,q1,q2e2),upper=c(q1,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe1
    te2d2opp[i]=(pmvnorm(lower=c(q1,-Inf,q2e1),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(q1,q1,q2e2),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe2
}

z = (1:300)/100.
par(mfrow=c(2,2))
plot(z,te1d1,ylim=c(0.,1),type="l")
plot(z,te1d2,ylim=c(0.,1),type="l")
plot(z,te2d1,ylim=c(0.,1),type="l")
plot(z,te2d2,ylim=c(0.,1.2),type="l")
plot(z,te1d1opp,ylim=c(0.,1),type="l")
plot(z,te1d2opp,ylim=c(0.,1),type="l")
plot(z,te2d1opp,ylim=c(0.,1),type="l")
plot(z,te2d2opp,ylim=c(0.,1.2),type="l")
```

lets do the simulation
```{r}
Y1<-sim2d(rv1,rhx1,b1,100,0.1,91,2,100)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.1,91,2,100)

for( z in 2:90){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z-1]*a
}

faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1]=1
faciesenv[Y1>q1 ]=2
both<-faciesenv
faciesdiag[]=2
faciesdiag[Y1<q1 & Z2<q2e1 ]=1
faciesdiag[Y1>q1 & Z2<q2e2 ]=1
both[faciesdiag==1]=3
image(faciesenv,col=c("white","grey"))
#image(faciesdiag)
image(both,col=c("white","grey","red"))
```

okay now that we hae studied some transiograms, the aim is to generate beautiful images with more facies

```{r}

pe1=0.4
pe2=0.4
pe3=0.2
pe1d1=0.3
pe2d1=0.3
pe3d1=0.3
rv1=0.3
rv2=0.3
rhx1=60
rhx2=60
b1=0
b2=0
a=0.8
shift=0.1
rho1a<-model(shift,rv1,b1)

q1a=qnorm(pe1) # threshold for environment of deposition
q1b=qnorm(pe1+pe2)

q2e1=q2e2=q2e3=0 # two thresholds for diagenetic facies 
Sigma = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
obj1=obj2=obj3=Inf
for ( test in 1:100 ){
  q2e1test = qnorm(test/100)
  q2e2test = qnorm(test/100)
  q2e3test = qnorm(test/100)
  newobj1 =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1a,q2e1test),mean=c(0,0),sigma=Sigma)/pe1 - pe1d1) 
  newobj2 =abs( pmvnorm(lower=c(q1a,-Inf),upper=c(q1b,q2e2test),mean=c(0,0),sigma=Sigma)/pe2 - pe2d1)
  newobj3 =abs( pmvnorm(lower=c(q1b,-Inf),upper=c(Inf,q2e3test),mean=c(0,0),sigma=Sigma)/pe3 - pe3d1)
  if (newobj1 < obj1){
    q2e1= q2e1test
    obj1 = newobj1
  }
  if (newobj2 < obj2){
    q2e2= q2e2test
    obj2 = newobj2
  }
  if (newobj3 < obj3){
    q2e3= q2e3test
    obj3 = newobj3
  }
}

Y1<-sim2d(rv1,rhx1,b1,100,0.05,91,1,100)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.05,91,1,100)

for( z in 2:90){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z-1]*a
}

faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1a]=1
faciesenv[Y1>q1a & Y1<q1b ]=2
faciesenv[Y1>q1b ]=3
combined<-faciesenv
faciesdiag[]=2
faciesdiag[Y1<q1a & Z2<q2e1 ]=1
faciesdiag[Y1>q1a & Y1<q1b & Z2<q2e2 ]=1
faciesdiag[Y1>q1b & Z2<q2e3 ]=1
combined[faciesenv==1 & faciesdiag==1]=4
combined[faciesenv==2 & faciesdiag==1]=5
combined[faciesenv==3 & faciesdiag==1]=6
image(faciesenv,col=c("ivory1","ivory3","ivory4"))
image(combined,col=c("ivory1","ivory3","ivory4","gold","darkorange","firebrick"))
```
same facies here but with vertical anisotropy for some so that it looks like fracture related diagenesis
```{r}
pe1=0.4
pe2=0.4
pe3=0.2
pe1d1=0.3
pe2d1=0.3
pe3d1=0.3
rv1=0.3
rv2=10
rhx1=60
rhx2=5
b1=0
b2=0
a=0.8
shift=0.1
rho1a<-model(shift,rv1,b1)

q1a=qnorm(pe1) # threshold for environment of deposition
q1b=qnorm(pe1+pe2)

q2e1=q2e2=q2e3=0 # two thresholds for diagenetic facies 
Sigma = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
obj1=obj2=obj3=Inf
for ( test in 1:100 ){
  q2e1test = qnorm(test/100)
  q2e2test = qnorm(test/100)
  q2e3test = qnorm(test/100)
  newobj1 =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1a,q2e1test),mean=c(0,0),sigma=Sigma)/pe1 - pe1d1) 
  newobj2 =abs( pmvnorm(lower=c(q1a,-Inf),upper=c(q1b,q2e2test),mean=c(0,0),sigma=Sigma)/pe2 - pe2d1)
  newobj3 =abs( pmvnorm(lower=c(q1b,-Inf),upper=c(Inf,q2e3test),mean=c(0,0),sigma=Sigma)/pe3 - pe3d1)
  if (newobj1 < obj1){
    q2e1= q2e1test
    obj1 = newobj1
  }
  if (newobj2 < obj2){
    q2e2= q2e2test
    obj2 = newobj2
  }
  if (newobj3 < obj3){
    q2e3= q2e3test
    obj3 = newobj3
  }
}

Y1<-sim2d(rv1,rhx1,b1,100,0.05,91,1,100)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.05,91,1,100)

for( z in 2:90){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z-1]*a
}

faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1a]=1
faciesenv[Y1>q1a & Y1<q1b ]=2
faciesenv[Y1>q1b ]=3
combined<-faciesenv
faciesdiag[]=2
faciesdiag[Y1<q1a & Z2<q2e1 ]=1
faciesdiag[Y1>q1a & Y1<q1b & Z2<q2e2 ]=1
faciesdiag[Y1>q1b & Z2<q2e3 ]=1
combined[faciesenv==1 & faciesdiag==1]=4
combined[faciesenv==2 & faciesdiag==1]=5
combined[faciesenv==3 & faciesdiag==1]=6
image(faciesenv,col=c("ivory1","ivory3","ivory4"))
image(combined,col=c("ivory1","ivory3","ivory4","gold","darkorange","firebrick"))
```

in three dimensions
```{r}
Z1<-sim3d(r1v,r1l,r1l,0,100,0.1,50,5,50,5,50)
Z2<-sim3d(r2v,r2l,r2l,0,100,0.1,50,5,50,5,50)
Y2<-Z2
for( z in 2:50){
  Z2[,,z]=Y2[,,z]*sqrt(1-(a^2))+Z1[,,z-1]*a
}
faciesenv<-faciesdiag<-Z1
faciesenv[Z1<q1a]=1
faciesenv[Z1>q1a & Z1<q1b ]=2
faciesenv[Z1>q1b ]=3
combined<-faciesenv
faciesdiag[]=2
faciesdiag[Z1<q1a & Z2<q2e1 ]=1
faciesdiag[Z1>q1a & Z1<q1b & Z2<q2e2 ]=1
faciesdiag[Z1>q1b & Z2<q2e3 ]=1
combined[faciesenv==1 & faciesdiag==1]=4
combined[faciesenv==2 & faciesdiag==1]=5
combined[faciesenv==3 & faciesdiag==1]=6

for ( z in 1:50){
  for ( y in 1:50){
    for ( x in 1:50){
      write(toString(c(combined[x,y,z])),"fracturecombined.csv",append=TRUE)
    }
  }
}
```

okay, we are going to change slighlty the simulation. 
We now do a pgs simulation for the environment of depositions so that we can have cyclicity. 
And the diagenesis happens only in two of the facies intertidal and supratidal. 
It might be better to visualize it like this.

```{r}
pe1=0.34
pe2=0.33
pe3=0.34
pe1d1=0.
pe2d1=0.3
pe3d1=0.4
rv1=0.2
rv2=0.2
rv3=0.2
rhx1=60
rhx2=60
rhx3=200
b1=0
b2=0
b3=0
a=0.99
b=0.99
shift=0.1
shiftb=-0.1
rho1a<-model(shift,rv1,b1)
rho1b<-model(shiftb,rv1,b1)

q1=qnorm(pe1) # subtidal facies defined by one Gaussian function

# for intertidal and supratidal, second Gaussian function
Sigma2 = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
q2=0 
obj = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q2test),mean=c(0,0),sigma=Sigma2) - pe3 ) 
  if (newobj < obj){
    q2 = q2test
    obj = newobj
  }
}

q3e1=-Inf
q3e2=q3e3=0 # two thresholds for diagenetic facies 
Sigma3 = diag(1,3)
Sigma3[1,2]=Sigma3[2,1]=a*rho1a
Sigma3[1,3]=Sigma3[3,1]=b*rho1b
Sigma3[2,3]=Sigma3[3,2]=a*b*model(shift-shiftb,rv1,b1)

obj1=obj2=obj3=Inf
for ( test in 1:1000 ){
  q3e2test = qnorm(test/1000)
  q3e3test = qnorm(test/1000)
  newobj2 =abs( pmvnorm(lower=c(q1,q2,-Inf),upper=c(Inf,Inf,q3e2test),mean=c(0,0,0),sigma=Sigma3)/pe2 - pe2d1)
  newobj3 =abs( pmvnorm(lower=c(q1,-Inf,-Inf),upper=c(Inf,q2,q3e3test),mean=c(0,0,0),sigma=Sigma3)/pe3 - pe3d1)
  if (newobj2 < obj2){
    q3e2= q3e2test
    obj2 = newobj2
  }
  if (newobj3 < obj3){
    q3e3= q3e3test
    obj3 = newobj3
  }
}

# Te2de2 as if there is only one facies field, which means that I will have a matrix slightly bigger.
# Z1(x)>q1 Z2(x)>q2 Z1(x+h)>q1 Z2(x+h)>q2 Z3(x+h)<q3e2 Z3(x)>q3e2
# Which means we have a 6*6 matrix
# 
# Cov[Z3(x),Z1(x)]=cov[bZ1(x+beta),Z1(x)]
# Cov[Z3(x),Z2(x)]=cov[bZ1(x+beta),aZ1(x+alpha)]
# 
# Cov[Z2(x),Z2(x+h)]=Cov[aZ1(x+alpha)+sqrt(1-a)Y2(x), aZ1(x+h+alpha)+sqrt(1-a)Y2(x+h)]

# #transiograms
te1de2=te2de2=te3de2=te1de3=te2de3=te3de3=rep(0,300)
te1de2opp=te2de2opp=te3de2opp=te1de3opp=te2de3opp=te3de3opp=rep(0,300)
Sigma = diag(1,6)
Sigmaopp = diag(1,6)
Sigma[1,2]=Sigma[2,1]=Sigma[3,4]=Sigma[4,3]=a*rho1a
Sigma[3,5]=Sigma[5,3]=Sigma[1,6]=Sigma[6,1]=b*rho1b
Sigma[4,5]=Sigma[5,4]=Sigma[2,6]=Sigma[6,2]=a*b*model(shiftb-shift,rv1,b1)

Sigmaopp = Sigma
for (i in 1:300){
    d = i/100.
    Sigma[1,3]=Sigma[3,1]= model(d,rv1,b1)
    Sigma[1,4]=Sigma[4,1]=a*model(d+shift,rv1,b1)
    Sigma[1,5]=Sigma[5,1]=b*model(d+shiftb,rv1,b1)
    Sigma[2,3]=Sigma[3,2]=a*model(d-shift,rv1,b1)
    Sigma[2,4]=Sigma[4,2]= (a^2)*model(d,rv1,b1)+sqrt(1-a^2)*model(d,rv2,b2)
    Sigma[2,5]=Sigma[5,2]=a*b*model(d+shiftb-shift,rv1,b1)

    Sigma[3,6]=Sigma[6,3]=b*model(d-shiftb,rv1,b1)
    Sigma[4,6]=Sigma[6,4]=a*b*model(d+shift-shiftb,rv1,b1)
    Sigma[5,6]=Sigma[6,5]=(b^2)*model(d,rv1,b1)+sqrt(1-b^2)*model(d,rv3,b3)

    Sigmaopp[1,3]=Sigmaopp[3,1]= model(-d,rv1,b1)
    Sigmaopp[1,4]=Sigmaopp[4,1]=a*model(-d+shift,rv1,b1)
    Sigmaopp[1,5]=Sigmaopp[5,1]=b*model(-d+shiftb,rv1,b1)
    Sigmaopp[2,3]=Sigmaopp[3,2]=a*model(-d-shift,rv1,b1)
    Sigmaopp[2,4]=Sigmaopp[4,2]= (a^2)*model(-d,rv1,b1)+sqrt(1-a^2)*model(-d,rv2,b2)
    Sigmaopp[2,5]=Sigmaopp[5,2]=a*b*model(-d+shiftb-shift,rv1,b1)

    Sigmaopp[3,6]=Sigmaopp[6,3]=b*model(-d-shiftb,rv1,b1)
    Sigmaopp[4,6]=Sigmaopp[6,4]=a*b*model(-d+shift-shiftb,rv1,b1)
    Sigmaopp[5,6]=Sigmaopp[6,5]=(b^2)*model(-d,rv1,b1)+sqrt(1-b^2)*model(-d,rv3,b3)

    te1de2[i]=pmvnorm(lower=c(-Inf,-Inf,q1,q2,-Inf,-Inf),upper=c(q1,Inf,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe1
    te2de2[i]=pmvnorm(lower=c(q1,q2,q1,q2,-Inf,q3e2),upper=c(Inf,Inf,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe2
    te3de2[i]=pmvnorm(lower=c(q1,-Inf,q1,q2,-Inf,q3e3),upper=c(Inf,q2,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe3

    te1de3[i]=pmvnorm(lower=c(-Inf,-Inf,q1,-Inf,-Inf,-Inf),upper=c(q1,Inf,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe1
    te2de3[i]=pmvnorm(lower=c(q1,q2,q1,-Inf,-Inf,q3e2),upper=c(Inf,Inf,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe2
    te3de3[i]=pmvnorm(lower=c(q1,-Inf,q1,-Inf,-Inf,q3e3),upper=c(Inf,q2,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigma)/ pe3

    te1de2opp[i]=pmvnorm(lower=c(-Inf,-Inf,q1,q2,-Inf,-Inf),upper=c(q1,Inf,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe1
    te2de2opp[i]=pmvnorm(lower=c(q1,q2,q1,q2,-Inf,q3e2),upper=c(Inf,Inf,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe2
    te3de2opp[i]=pmvnorm(lower=c(q1,-Inf,q1,q2,-Inf,q3e3),upper=c(Inf,q2,Inf,Inf,q3e2,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe3

    te1de3opp[i]=pmvnorm(lower=c(-Inf,-Inf,q1,-Inf,-Inf,-Inf),upper=c(q1,Inf,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe1
    te2de3opp[i]=pmvnorm(lower=c(q1,q2,q1,-Inf,-Inf,q3e2),upper=c(Inf,Inf,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe2
    te3de3opp[i]=pmvnorm(lower=c(q1,-Inf,q1,-Inf,-Inf,q3e3),upper=c(Inf,q2,Inf,q2,q3e3,Inf),mean=c(0,0,0,0,0,0),sigma=Sigmaopp)/ pe3

    # te1d1opp[i]=(pmvnorm(lower=c(-Inf,-Inf,-Inf),upper=c(q1,q1,q2e1),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(-Inf,q1,-Inf),upper=c(q1,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe1
    # te2d1opp[i]=(pmvnorm(lower=c(q1,-Inf,-Inf),upper=c(Inf,q1,q2e1),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(q1,q1,-Inf),upper=c(Inf,Inf,q2e2),mean=c(0,0,0),sigma=Sigma))/ pe2
    # te1d2opp[i]=(pmvnorm(lower=c(-Inf,-Inf,q2e1),upper=c(q1,q1,Inf),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(-Inf,q1,q2e2),upper=c(q1,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe1
    # te2d2opp[i]=(pmvnorm(lower=c(q1,-Inf,q2e1),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=Sigmaopp) + pmvnorm(lower=c(q1,q1,q2e2),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=Sigma))/ pe2
}
z = (1:300)/100.
par(mfrow=c(2,3))
plot(z,te1de2,ylim=c(0.,1),type="l")
plot(z,te2de2,ylim=c(0.,1),type="l")
plot(z,te3de2,ylim=c(0.,1),type="l")
plot(z,te1de3,ylim=c(0.,1),type="l")
plot(z,te2de3,ylim=c(0.,1),type="l")
plot(z,te3de3,ylim=c(0.,1),type="l")

z = (1:300)/100.
par(mfrow=c(2,3))
plot(z,te1de2opp,ylim=c(0.,1),type="l")
plot(z,te2de2opp,ylim=c(0.,1),type="l")
plot(z,te3de2opp,ylim=c(0.,1),type="l")
plot(z,te1de3opp,ylim=c(0.,1),type="l")
plot(z,te2de3opp,ylim=c(0.,1),type="l")
plot(z,te3de3opp,ylim=c(0.,1),type="l")

#simulation
Y1<-sim2d(rv1,rhx1,b1,100,0.025,182,0.5,200)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.025,182,0.5,200)
Z3<-Y3<-sim2d(rv3,rhx3,b3,100,0.025,182,0.5,200)
for( z in 5:178){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z+4]*a
}
for( z in 5:178){
  Z3[,z]=Y3[,z]*sqrt(1-(b^2))+Y1[,z-4]*b
}
faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1]=1
faciesenv[Y1>q1 & Z2<q2 ]=3
faciesenv[Y1>q1 & Z2>q2  ]=2
combined<-faciesenv
faciesdiag[]=2
faciesdiag[Y1>q1 & Z2<q2 & Z3<q3e3 ]=1
faciesdiag[Y1>q1 & Z2>q2 & Z3<q3e2 ]=1
combined[faciesenv==2 & faciesdiag==1]=4
combined[faciesenv==3 & faciesdiag==1]=5
par(mfrow=c(1,1))
image(faciesenv,col=c("ivory1","ivory3","ivory4"))
image(combined,col=c("ivory1","ivory3","ivory4","gold","firebrick"),axes = FALSE)
#image(combined,col=c("ivory1","ivory3","ivory4","gold"),axes = FALSE)
#image(combined,col=c("ivory1","ivory3","ivory4","firebrick"),axes = FALSE)
```
Okay, we can do this simulation in three dimensions
```{r}
Y1<-sim3d(rv1,rhx1,rhx1,b1,100,0.025,182,0.5,200,5,10)
Z2<-Y2<-sim3d(rv2,rhx2,rhx2,b2,100,0.025,182,0.5,200,5,10)
Z3<-Y3<-sim3d(rv3,rhx3,rhx3,b3,100,0.025,182,0.5,200,5,10)
for( z in 5:178){
  Z2[,,z]=Y2[,,z]*sqrt(1-(a^2))+Y1[,,z+4]*a
}
for( z in 5:178){
  Z3[,,z]=Y3[,,z]*sqrt(1-(b^2))+Y1[,,z-4]*b
}

```

```{r}
faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1]=1
faciesenv[Y1>q1 & Z2<q2 ]=3
faciesenv[Y1>q1 & Z2>q2  ]=2
faciesenv[faciesenv==3 & Z3<q3e3]=4
faciesenv[faciesenv==2 & Z3<q3e2]=5
for ( z in 1:182){
  for ( y in 1:10){
    for ( x in 1:200){
      write(toString(c(faciesenv[x,y,z])),"fracturecombined.csv",append=TRUE)
    }
  }
}
```

here with vertical fractures dolomitized 
```{r}
pe1=0.34
pe2=0.33
pe3=0.34
pe1d1=0.
pe2d1=0.2
pe3d1=0.4
rv1=0.2
rv2=0.2
rv3=5
rhx1=60
rhx2=60
rhx3=5
b1=0
b2=0
b3=0
a=0.99
b=0.8
shift=0.1
shiftb=0.
rho1a<-model(shift,rv1,b1)
rho1b<-model(shiftb,rv1,b1)

q1=qnorm(pe1) # subtidal facies defined by one Gaussian function

# for intertidal and supratidal, second Gaussian function
Sigma2 = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
q2=0 
obj = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q2test),mean=c(0,0),sigma=Sigma2) - pe3 ) 
  if (newobj < obj){
    q2 = q2test
    obj = newobj
  }
}

q3e1=-Inf
q3e2=q3e3=0 # two thresholds for diagenetic facies 
Sigma3 = diag(1,3)
Sigma3[1,2]=Sigma3[2,1]=a*rho1a
Sigma3[1,3]=Sigma3[3,1]=b*rho1b
Sigma3[2,3]=Sigma3[3,2]=a*b*model(shift-shiftb,rv1,b1)

obj1=obj2=obj3=Inf
for ( test in 1:1000 ){
  q3e2test = qnorm(test/1000)
  q3e3test = qnorm(test/1000)
  newobj2 =abs( pmvnorm(lower=c(q1,q2,-Inf),upper=c(Inf,Inf,q3e2test),mean=c(0,0,0),sigma=Sigma3)/pe2 - pe2d1)
  newobj3 =abs( pmvnorm(lower=c(q1,-Inf,-Inf),upper=c(Inf,q2,q3e3test),mean=c(0,0,0),sigma=Sigma3)/pe3 - pe3d1)
  if (newobj2 < obj2){
    q3e2= q3e2test
    obj2 = newobj2
  }
  if (newobj3 < obj3){
    q3e3= q3e3test
    obj3 = newobj3
  }
}

Y1<-sim2d(rv1,rhx1,b1,100,0.025,182,0.5,200)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.025,182,0.5,200)
Z3<-Y3<-sim2d(rv3,rhx3,b3,100,0.025,182,0.5,200)
for( z in 4:178){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z+4]*a
}
for( z in 4:178){
  Z3[,z]=Y3[,z]*sqrt(1-(b^2))+Y1[,z]*b
}
faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1]=1
faciesenv[Y1>q1 & Z2<q2 ]=3
faciesenv[Y1>q1 & Z2>q2  ]=2
combined<-faciesenv
faciesdiag[]=2
faciesdiag[Y1>q1 & Z2<q2 & Z3<q3e3 ]=1
faciesdiag[Y1>q1 & Z2>q2 & Z3<q3e2 ]=1
combined[faciesenv==2 & faciesdiag==1]=4
combined[faciesenv==3 & faciesdiag==1]=5
image(faciesenv,col=c("ivory1","ivory3","ivory4"))
image(combined,col=c("ivory1","ivory3","ivory4","gold","firebrick"),axes = FALSE)
```
in three dimensions
```{r}
Y1<-sim3d(rv1,rhx1,rhx1,b1,100,0.025,182,0.5,200,5,10)
Z2<-Y2<-sim3d(rv2,rhx2,rhx2,b2,100,0.025,182,0.5,200,5,10)
Z3<-Y3<-sim3d(rv3,rhx3,rhx3,b3,100,0.025,182,0.5,200,5,10)
for( z in 4:178){
  Z2[,,z]=Y2[,,z]*sqrt(1-(a^2))+Y1[,,z+1]*a
}
for( z in 4:178){
  Z3[,,z]=Y3[,,z]*sqrt(1-(b^2))+Y1[,,z]*b
}

```

```{r}
faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1]=1
faciesenv[Y1>q1 & Z2<q2 ]=3
faciesenv[Y1>q1 & Z2>q2  ]=2
faciesenv[faciesenv==3 & Z3<q3e3]=4
faciesenv[faciesenv==2 & Z3<q3e2]=5
for ( z in 1:182){
  for ( y in 1:10){
    for ( x in 1:200){
      write(toString(c(faciesenv[x,y,z])),"fracturecombined.csv",append=TRUE)
    }
  }
}
```

new simulation with four facies with the first two Gaussian functions
The fourth facies is marine cementation during transgression phase. This facies is the last of the cycle.
The diagenesic facies is dolomite and just happen in the supratidal environment.

```{r}
pe1=0.2
pe2=0.2
pe3=0.2
pe4=0.2
pe3d1=0.4
rv1=0.2
rv2=0.2
rv3=0.2
rhx1=60
rhx2=60
rhx3=60
b1=0
b2=0
b3=0
a=0.99
b=0.
shift=0.1
shiftb=0.
rho1a<-model(shift,rv1,b1)
rho1b<-model(shiftb,rv1,b1)

q1=qnorm(pe1+pe2) # subtidal facies defined by one Gaussian function

# for intertidal and supratidal, second Gaussian function
Sigma2 = matrix(c(1,a*rho1a,a*rho1a,1),2,2)
q2=0
q3=0
obj=obj2 = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  q3test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(-Inf,-Inf),upper=c(q1,q2test),mean=c(0,0),sigma=Sigma2) - pe1 ) 
  newobj2 =abs( pmvnorm(lower=c(q1,-Inf),upper=c(Inf,q3test),mean=c(0,0),sigma=Sigma2) - pe4 )
  if (newobj < obj){
    q2 = q2test
    obj = newobj
  }
  if (newobj2 < obj2){
    q3 = q3test
    obj2 = newobj2
  }
}

q3e3=-Inf
Sigma3 = diag(1,3)
Sigma3[1,2]=Sigma3[2,1]=a*rho1a
Sigma3[1,3]=Sigma3[3,1]=b*rho1b
Sigma3[2,3]=Sigma3[3,2]=a*b*model(shift-shiftb,rv1,b1)

obj3=Inf
for ( test in 1:1000 ){
  q3e3test = qnorm(test/1000)
  newobj3 =abs( pmvnorm(lower=c(q1,q2,-Inf),upper=c(Inf,Inf,q3e3test),mean=c(0,0,0),sigma=Sigma3)/pe3 - pe3d1)

  if (newobj3 < obj3){
    q3e3= q3e3test
    obj3 = newobj3
  }
}

#simulation
Y1<-sim2d(rv1,rhx1,b1,100,0.025,182,0.5,200)
Z2<-Y2<-sim2d(rv2,rhx2,b2,100,0.025,182,0.5,200)
Z3<-Y3<-sim2d(rv3,rhx3,b3,100,0.025,182,0.5,200)
for( z in 4:178){
  Z2[,z]=Y2[,z]*sqrt(1-(a^2))+Y1[,z+4]*a
}
for( z in 4:178){
  Z3[,z]=Y3[,z]*sqrt(1-(b^2))+Y1[,z]*b
}
faciesenv<-faciesdiag<-Y1
faciesenv[Y1<q1 & Z2<q2]=1
faciesenv[Y1<q1 & Z2>q2]=2
faciesenv[Y1>q1 & Z2>q3]=3
faciesenv[Y1>q1 & Z2<q3]=4

combined<-faciesenv
combined[faciesenv==3 & Z3<q3e3]=5
image(faciesenv,col=c("ivory1","ivory3","ivory4"))
image(combined,col=c("ivory1","ivory3","ivory4","cyan3","firebrick"),axes = FALSE)
```