### 2018-05-31 cosim_expo.Rmd

---
title: "R Notebook"
output: html_notebook
---

we have seen that the conditioning was difficult with a Gaussian covariance. 
Therefore, we perform the same method that precedently, with exponential covariances. 
We have proved that the covariance matrix was more stable, and could be inverted safely. 
We may still use a Gaussian covariance in the lateral direction, it is more stable because less data in this direction.

```{r}
source("cosim_functions.R")
```
```{r}
library(mvtnorm)
```

```{r}
model<-function(h,range,b){return(exp(-abs(h/range))*cos(b*h) )}
model_lat<-function(h,range){return(exp(-(h/range)^2) )}
```

```{r}
rv1=0.3
rhx1=50
b1=0
Z3<-sim2dexp(rv3,rhx3,b3,100,0.1,91,1,100)
image(Z3,col = heat.colors(100))
```

```{r}
rv1=0.3
rv2=0.5
rhx1=100
rhx2=100
b1=0
b2=0
rho=0.6
shift=0.1
rho1a<-model(shift,rv1,b1)

alpha1=(rho/rho1a)^2
alpha2=1-(rho/rho1a)^2
rv3=0.2
rv4=0.3
rhx3=100
rhx4=100
b3=0
b4=0
rho34=0.
c=0.
e=0.65
a13=-0.
a14=-0.1
rho1a14 = model(a14,rv1,b1)
```

thresholds of diagenetic facies, depend on the parameters

```{r}
qd1e4=qnorm(pd3e4)

# now the threshold with the fourth Gaussian function, that does depend on the first
sigma1=diag(1,3)
sigma1[1,2]=sigma1[2,1]=e*model(a14,rv1,b1)
sigma1[1,3]=sigma1[3,1]=e*(rho/rho1a)*model(a14-shift,rv1,b1)
sigma1[2,3]=sigma1[3,2]=rho

qd2e2=0
obj = Inf
for ( test in 1:100 ){
  qd2e2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(qd2e2test,-Inf,q2),upper=c(Inf,q1,Inf),mean=c(0,0,0),sigma=sigma1)/pe2 - pd2e2 ) 
  if (newobj < obj){
    qd2e2 = qd2e2test
    obj = newobj
  }
}

qd2e3=0
obj = Inf
for ( test in 1:100 ){
  qd2e3test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(qd2e3test,q1,q3),upper=c(Inf,Inf,Inf),mean=c(0,0,0),sigma=sigma1)/pe3 - pd2e3 ) 
  if (newobj < obj){
    qd2e3 = qd2e3test
    obj = newobj
  }
}

print(c(qd1e4,qd2e2,qd2e3))
```

simu

```{r}
Z1<-sim2dexp(rv1,rhx1,b1,100,0.05,182,0.5,200)
Y2<-sim2dexp(rv2,rhx2,b2,100,0.05,182,0.5,200)
Z2=Y2
for ( z in 1:180){
  Z2[,z]=Y2[,z]*sqrt(1-(rho^2)/(rho1a^2))+Z1[,z+2]*(rho)/(rho1a)
}
Z3<-sim2dexp(rv3,rhx3,b3,100,0.05,182,0.5,200)
Y4<-sim2dexp(rv4,rhx4,b4,100,0.05,182,0.5,200)
Z4=Y4
for ( z in 2:182){
  Z4[,z]=Y4[,z]*sqrt(1-e^2)+Z1[,z-1]*e
}

faciesenv<-faciesdiag<-Z1
faciesenv[Z1<q1 & Z2<q2]=1
faciesenv[Z1<q1 & Z2>q2]=2
faciesenv[Z1>q1 & Z2<q3]=4
faciesenv[Z1>q1 & Z2>q3]=3
faciesdiag[]=1
faciesdiag[Z1>q1 & Z2<q3 & Z3<qd1e4]=3
faciesdiag[Z1<q1 & Z2>q2 & Z4>qd2e2]=2
faciesdiag[Z1>q1 & Z2>q3 & Z4>qd2e3]=2
col_env <- c("lightgreen","cadetblue","dodgerblue4","lightcoral")
col_diag<-dcolors <- c("black","grey","yellow")
image(faciesenv,col = col_env)
image(faciesdiag,col = col_diag)
bin<-bin2<-faciesenv
#bin[]<-rbinom(91*100, 1, 0.5)
bin[]=bin2[]=0
bin[seq(2,182*200,by=2)]=1
facies=faciesenv
facies[faciesdiag==2 & bin ==1 ] = 5
facies[faciesdiag==3]=5

col_combine <- c("lightgreen","cadetblue","dodgerblue4","lightcoral","yellow")
W<-image(facies,col = col_combine)
```

we need to look again at the experimental transiogram, to do an accurate fitting with the exponential covariance.
transiograms of environment of depositions

```{r}
rv1=0.6
rv2=0.7
rhx1=50
rhx2=50
b1=5
b2=0
rho=0.7
shift=0.1
rho1a<-model(shift,rv1,b1)

alpha1=(rho/rho1a)^2
alpha2=1-(rho/rho1a)^2
rv3=0.4
rv4=0.3
rhx3=50
rhx4=50
b3=0
b4=0
rho34=0.
c=0.
e=0.65
a13=-0.
a14=-0.1
rho1a14 = model(a14,rv1,b1)

z = (1:300)/100.
t11_model=t12_model=t13_model=t21_model=t22_model=t23_model=t31_model=t32_model=t33_model=rep(0,300)
Sigma = diag(1,4)
for (i in 1:300){
    d = i/100.
    Sigma[1,2]=Sigma[2,1]= Sigma[3,4]=Sigma[4,3]=rho
    Sigma[1,3]=Sigma[3,1]=model(d,rv1,b1)
    Sigma[1,4]=Sigma[4,1]=(rho/rho1a)*model(abs(d+shift),rv1,b1)
    Sigma[2,3]=Sigma[3,2]=(rho/rho1a)*model(abs(d-shift),rv1,b1)
    Sigma[2,4]=Sigma[4,2]=alpha1*model(d,rv1,b1)+alpha2*model(d,rv2,b2)
    t11_model[i]=pmvnorm(lower=c(-Inf,-Inf,-Inf,-Inf),upper=c(q1,q2,q1,q2),mean=c(0,0,0,0),sigma=Sigma)/ pe1 
    t12_model[i]=pmvnorm(lower=c(-Inf,-Inf,-Inf,q2),upper=c(q1,q2,q1,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe1 
    t13_model[i]=pmvnorm(lower=c(-Inf,-Inf,q1,q3),upper=c(q1,q2,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe1 
    t21_model[i]=pmvnorm(lower=c(-Inf,q2,-Inf,-Inf),upper=c(q1,Inf,q1,q2),mean=c(0,0,0,0),sigma=Sigma)/ pe2
    t22_model[i]=pmvnorm(lower=c(-Inf,q2,-Inf,q2),upper=c(q1,Inf,q1,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe2
    t23_model[i]=pmvnorm(lower=c(-Inf,q2,q1,q3),upper=c(q1,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe2
    t31_model[i]=pmvnorm(lower=c(q1,q3,-Inf,-Inf),upper=c(Inf,Inf,q1,q2),mean=c(0,0,0,0),sigma=Sigma)/ pe3
    t32_model[i]=pmvnorm(lower=c(q1,q3,-Inf,q2),upper=c(Inf,Inf,q1,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe3
    t33_model[i]=pmvnorm(lower=c(q1,q3,q1,q3),upper=c(Inf,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pe3
}
par(mfrow=c(3,3))
t11<-tr_2d_vert(e1,e1,30,nz,0.1)
lines(z,t11_model)
t12<-tr_2d_vert(e1,e2,30,nz,0.1)
lines(z,t12_model)
t13<-tr_2d_vert(e1,e3,30,nz,0.1)
lines(z,t13_model)
t21<-tr_2d_vert(e2,e1,30,nz,0.1)
lines(z,t21_model)
t22<-tr_2d_vert(e2,e2,30,nz,0.1)
lines(z,t22_model)
t23<-tr_2d_vert(e2,e3,30,nz,0.1)
lines(z,t23_model)
t31<-tr_2d_vert(e3,e1,30,nz,0.1)
lines(z,t31_model)
t32<-tr_2d_vert(e3,e2,30,nz,0.1)
lines(z,t32_model)
t33<-tr_2d_vert(e3,e3,30,nz,0.1)
lines(z,t33_model)
```

now the diagenetic transiograms

```{r}

rv3=0.2
rv4=0.6
qd1=qnorm(pd3)
Sigma1 = matrix(c(1,rho34,rho34,1),2,2)
qd2=0
obj = Inf
for ( test in 1:100 ){
  q2test = qnorm(test/100)
  newobj =abs( pmvnorm(lower=c(qd1,q2test),upper=c(Inf,Inf),mean=c(0,0),sigma=Sigma1) - pd2 ) 
  if (newobj < obj){
    qd2 = q2test
    obj = newobj
  }
}

c(qd1,qd2)

#transiograms

z = (1:300)/100.
t33d_model=t32d_model=t23d_model=t22d_model=rep(0,300)
Sigma = diag(1,4)
for (i in 1:300){
    d = i/100.
    Sigma[1,2]=Sigma[2,1]= Sigma[3,4]=Sigma[4,3]=rho34
    Sigma[1,3]=Sigma[3,1]=model(d,rv3,b3)
    Sigma[1,4]=Sigma[4,1]=Sigma[2,3]=Sigma[3,2]=rho34*model(d,rv3,b3)
    Sigma[2,4]=Sigma[4,2]=(rho34^2)*model(d,rv3,b3)+(1-rho34^2)*model(d,rv4,b4)

    t22d_model[i]=pmvnorm(lower=c(qd1,qd2,qd1,qd2),upper=c(Inf,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pd2
    t23d_model[i]=pmvnorm(lower=c(qd1,qd2,-Inf,-Inf),upper=c(Inf,Inf,qd1,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pd2
    t32d_model[i]=pmvnorm(lower=c(-Inf,-Inf,qd1,qd2),upper=c(qd1,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pd3
    t33d_model[i]=pmvnorm(lower=c(-Inf,-Inf,-Inf,-Inf),upper=c(qd1,Inf,qd1,Inf),mean=c(0,0,0,0),sigma=Sigma)/ pd3
}
par(mfrow=c(2,2))
t22d<-tr_2d_vert(d2,d2,30,nz,0.1)
lines(z,t22d_model)
t23d<-tr_2d_vert(d2,d3,30,nz,0.1)
lines(z,t23d_model)
t32d<-tr_2d_vert(d3,d2,30,nz,0.1)
lines(z,t32d_model)
t33d<-tr_2d_vert(d3,d3,30,nz,0.1)
lines(z,t33d_model)
```

conditioning

```{r}
source("co_sim_condi.R")
```

```{r}
rv1=0.6
rv2=0.7
rhx1=40
rhx2=40
#b1=5
b1=0
b2=0
rho=0.7
shift=0.1
rho1a<-model(shift,rv1,b1)

alpha1=(rho/rho1a)^2
alpha2=1-(rho/rho1a)^2
rv3=0.2
rv4=0.6
rhx3=40
rhx4=40
b3=0
b4=0
rho34=0.
c=0.
e=0.65
a13=-0.
a14=-0.1
rho1a14 = model(a14,rv1,b1)

Z1<-sim2dexp(rv1,rhx1,b1,100,0.1,91,1,100)
Y2<-sim2dexp(rv2,rhx2,b2,100,0.1,91,1,100)
Z2=Y2
for ( z in 1:90){
  Z2[,z]=Y2[,z]*sqrt(1-(rho^2)/(rho1a^2))+Z1[,z+1]*(rho)/(rho1a)
}
Z3<-sim2dexp(rv3,rhx3,b3,100,0.1,91,1,100)
Y4<-sim2dexp(rv4,rhx4,b4,100,0.1,91,1,100)
Z4=Y4
for ( z in 2:91){
  Z4[,z]=Y4[,z]*sqrt(1-e^2)+Z1[,z-1]*e
}
grid<-grid_co<-array(NA,c(100,91))
grid[3,]<-env_lat[,1]
grid[20,]<-env_lat[,2]
grid[55,]<-env_lat[,3]
grid[80,]<-env_lat[,4]
grid_co[3,]<-diag_lat[,1]
grid_co[20,]<-diag_lat[,2]
grid_co[55,]<-diag_lat[,3]
grid_co[80,]<-diag_lat[,4]

simwell<-coco_gibbs_sampling(grid,grid_co,c(q1,q2,q3),c(qd1e4,qd2e2,qd2e3),0.1,1,c(rv1,rv2,rv3,rv4),c(rhx1,rhx2,rhx3,rhx4),c(b1,b2,b3,b4),c(rho,e*model(a14,rv1,b1)),c(shift,a14),30)
Z1well<-simwell[,1:91]
Z2well<-simwell[,92:182]
Z3well<-simwell[,183:273]
Z4well<-simwell[,274:364]
E<-coco_dual_krig(Z1well-Z1,Z2well-Z2,Z3well-Z3,Z4well-Z4,0.1,1,c(rv1,rv2,rv3,rv4),c(rhx1,rhx2,rhx3,rhx4),c(b1,b2,b3,b4),c(rho,e*model(a14,rv1,b1)),c(shift,a14))
Z1c=Z1 + E[,1:91]
Z2c<-Z2 +E[,92:182]
Z3c<-Z3 +E[,183:273]
Z4c<-Z4 +E[,274:364]

faciesenv<-faciesdiag<-Z1c
faciesenv[Z1c<q1 & Z2c<q2]=1
faciesenv[Z1c<q1 & Z2c>q2]=2
faciesenv[Z1c>q1 & Z2c<q3]=4
faciesenv[Z1c>q1 & Z2c>q3]=3
faciesdiag[]=1
faciesdiag[Z1c>q1 & Z2c<q3 & Z3c<qd1e4]=3
faciesdiag[Z1c<q1 & Z2c>q2 & Z4c>qd2e2]=2
faciesdiag[Z1c>q1 & Z2c>q3 & Z4c>qd2e3]=2
col_env <- c("lightgreen","cadetblue","dodgerblue4","lightcoral")
W<-image(faciesenv,col = col_env)
bin<-bin2<-faciesenv
#bin[]<-rbinom(91*100, 1, 0.5)
bin[]=bin2[]=0
bin[seq(2,91*100,by=2)]=1
facies=faciesenv
facies[faciesdiag==2 & bin ==1 ] = 5
facies[faciesdiag==3]=5

col_combine <- c("lightgreen","cadetblue","dodgerblue4","lightcoral","yellow")
W<-image(facies,col = col_combine)
image(Z2c)
image(Z2)
```