### 2018-10-15 main_bookcliff.Rmd
cleaning

---
title: "R Notebook"
output: html_notebook
---
facies1=distal lower shoreface heteroliths & offshore mudstones
facies2=proximal lower shoreface sandstones
facies3=foreshore & upper shoreface sandstones

This case study is not stationary so we are importing functions
redefined in a non stationary case
```{r}
source("nonstation_functions.R")
```
```{r}
library(mvtnorm)
```

```{r}
i1=i2=i3=env_lat<-data.matrix(read.table("envclastic.txt"))
i1[env_lat!=1]=0
i2[env_lat==2]=1
i2[env_lat!=2]=0
i3[env_lat==3]=1
i3[env_lat!=3]=0
p1=mean(i1,na.rm=TRUE)
p2=mean(i2,na.rm=TRUE)
p3=mean(i3,na.rm=TRUE)
c(p1,p2,p3)
image(t(env_lat),col=c("blue","green","red"))
```

```{r}
par(mfrow=c(3,3))
t11exp=tr_2d_vert(i1,i1,200,566,56.6)
t12exp=tr_2d_vert(i1,i2,200,566,56.6)
t13exp=tr_2d_vert(i1,i3,200,566,56.6)
t21exp=tr_2d_vert(i2,i1,200,566,56.6)
t22exp=tr_2d_vert(i2,i2,200,566,56.6)
t23exp=tr_2d_vert(i2,i3,200,566,56.6)
t31exp=tr_2d_vert(i3,i1,200,566,56.6)
t32exp=tr_2d_vert(i3,i2,200,566,56.6)
t33exp=tr_2d_vert(i3,i3,200,566,56.6)
```

if we want to fit these transiograms, we need to implement non stationarity
therefore, we need to compute the proportion on each horizontal 
```{r}
prop1=prop2=prop3=numeric(nrow(env_lat))
for (i in 1:nrow(env_lat)){
  prop1[i]=mean(i1[i,],na.rm=TRUE)
  prop2[i]=mean(i2[i,],na.rm=TRUE)
  prop3[i]=mean(i3[i,],na.rm=TRUE)
}
plot(1:nrow(i1),prop1,type = "l")
plot(1:nrow(i1),prop2,type = "l")
plot(1:nrow(i1),prop3,type = "l")
```
ok we are now smoothing the curves with moving average
```{r}
f50=rep(1/50,50)
prop1filter<-filter(prop1,f50,sides=2)
prop1filter[542:566]=prop1[542:566]
prop1filter[1:24]=prop1[1:24]
plot(1:nrow(i1),prop1filter,type = "l")

prop3filter<-filter(prop3,f50,sides=2)
prop3filter[542:566]=prop3[542:566]
prop3filter[1:24]=prop3[1:24]
plot(1:nrow(i1),prop3filter,type = "l")
prop2filter=1-prop1filter-prop3filter
plot(1:nrow(i1),prop2filter,type = "l")

```
we want to plot an image with those curves
```{r}
png(file="PCvert.png",width=2000,height = 800,res=300,antialias = )
par(mar=c(0,0,0,0))
n=nrow(i1)
plot(1:n,prop2filter+prop3filter,type="l",axes=FALSE,ann=FALSE)
lines(1:n,prop3filter)
lines(1:n,rep(1,n))
lines(1:n,rep(0,n))
```

ok now we need to compute the theoretical transiograms with the accurate proportions
this is not good because the global proportions also depend on the horizontal PC

```{r}
model<-function(h,range,b){return(exp(-abs(h/range)^2)*cos(b*h) )}
```

okay, if we want to fit these transiograms, we need to implement non stationarity
therefore, we need to compute the proportion on each vertical 
```{r}
prop1lat=prop2lat=prop3lat=numeric(ncol(env_lat))
for (i in 1:ncol(env_lat)){
  prop1lat[i]=mean(i1[,i],na.rm=TRUE)
  prop2lat[i]=mean(i2[,i],na.rm=TRUE)
  prop3lat[i]=mean(i3[,i],na.rm=TRUE)
}
plot(1:ncol(i1),prop1lat,type = "p")
plot(1:ncol(i1),prop2lat,type = "p")
plot(1:ncol(i1),prop3lat,type = "p")
```
now doing a good figure
```{r}
png(file="PChor.png",width=2000,height = 800,res=300,antialias = )
par(mar=c(0,0,0,0))
n=ncol(i1)
plot(1:n,prop2lat+prop3lat,type="p",axes=FALSE,ann=FALSE,ylim=c(0,1))
lines(1:n,prop3lat,type="p")
lines(1:n,rep(1,n))
lines(1:n,rep(0,n))
```

we are now doing the mean of the lateral and vertical proportion curves
```{r}
prop1mat=prop2mat=prop3mat=matrix(0,nrow=nrow(env_lat),ncol=ncol(env_lat))
for ( i in 1:nrow(env_lat)){
  for (j in 1:ncol(env_lat)){
    prop1mat[i,j]=(prop1filter[i]+prop1lat[j])/2
    prop2mat[i,j]=(prop2filter[i]+prop2lat[j])/2
    #prop3mat[i,j]=(prop3filter[i]+prop3lat[j])/2
  }
}
prop3mat=1-prop2mat-prop1mat
prop3mat[prop3mat<0]=0
image(prop1mat)
image(prop2mat)
image(prop3mat)
```
Now that we have the proportions on the grid
we can compute the corresponding thresholds of the Gaussian functions according to the truncation rule

We can compute the thresholds analytically if the Gaussian functions are not correlated to each other here
```{r}
n=nrow(i1)
m=ncol(i1)
q1=q2=matrix(0,n,m)
for (z in 1:n){
  for(x in 1:m){
      q1[z,x]=qnorm(prop1mat[z,x])
      if (prop3mat[z,x]!=0) {
      q2[z,x]=qnorm(1-(prop2mat[z,x]/(1-prop1mat[z,x])))
      }
      else{q2[z,x]=-Inf}
  }
}
```
Because of the non stationarity, it is more relevant to computed the transition rates between facies rather than the
transiograms
we can thus fit the method with the transition rates

thresholds according to the truncation rule
```{r}
q1g=qnorm(p1)
q2g=qnorm(1-(p2/(1-p1)))
```

```{r}
rv1=5
rv2=5
rh=800
b1=0
b2=0
rho=0
rv1=0.6
rv2=0.7
Sigma = diag(1,4)
d = 1/10
Sigma[1,2]=Sigma[2,1]= Sigma[3,4]=Sigma[4,3]=rho
Sigma[1,3]=Sigma[3,1]=model(d,rv1,b1)
Sigma[1,4]=Sigma[4,1]=(rho)*model(abs(d),rv1,b1)
Sigma[2,3]=Sigma[3,2]=(rho)*model(abs(d),rv1,b1)
Sigma[2,4]=Sigma[4,2]=(rho^2)*model(d,rv1,b1)+(1-(rho^2))*model(d,rv2,b2)
r11_model=(1-pmvnorm(lower=c(-Inf,-Inf,-Inf,-Inf),upper=c(q1g,Inf,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p1)
r12_model=pmvnorm(lower=c(-Inf,-Inf,q1g,q2g),upper=c(q1g,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/(p1)
r13_model=pmvnorm(lower=c(-Inf,-Inf,q1g,-Inf),upper=c(q1g,Inf,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p1
r21_model=pmvnorm(lower=c(q1g,q2g,-Inf,-Inf),upper=c(Inf,Inf,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p2
r22_model=(1-pmvnorm(lower=c(q1g,q2g,q1g,q2g),upper=c(Inf,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/p2)
r23_model=pmvnorm(lower=c(q1g,q2g,q1g,-Inf),upper=c(Inf,Inf,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p2
r31_model=pmvnorm(lower=c(q1g,-Inf,-Inf,-Inf),upper=c(Inf,q2g,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p3
r32_model=pmvnorm(lower=c(q1g,-Inf,q1g,q2g),upper=c(Inf,q2g,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/p3
r33_model=(1-pmvnorm(lower=c(q1g,-Inf,q1g,-Inf),upper=c(Inf,q2g,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p3)
matrix(c(0,r12_model/r11_model,r13_model/r11_model,r21_model/r22_model,0,r23_model/r22_model,r31_model/r33_model,r32_model/r33_model,0),3,3,byrow=TRUE)

```
lets see what the thicknesses are
```{r}
(1/r11_model)*0.1
(1/r22_model)*0.1
(1/r33_model)*0.1
```

```{r}
r11exp=-(1-t11exp[1])/0.1
r12exp=t12exp[1]/0.1
r13exp=t13exp[1]/0.1
r21exp=t21exp[1]/0.1
r22exp=-(1-t22exp[1])/0.1
r23exp=t23exp[1]/0.1
r31exp=t31exp[1]/0.1
r32exp=t32exp[1]/0.1
r33exp=-(1-t33exp[1])/0.1
matrix(c(r11exp,r12exp,r13exp,r21exp,r22exp,r23exp,r31exp,r32exp,r33exp),3,3)
```
and mean thicknesses
```{r}
-(1/r11exp)*0.1
-(1/r22exp)*0.1
-(1/r33exp)*0.1
```

okay now we create a grid (3d) to simulate the facies

```{r}
grid<-array(NA,c(100,1,566))
grid[5,1,]<-env_lat[,1]
grid[15,1,]<-env_lat[,2]
grid[25,1,]<-env_lat[,3]
grid[35,1,]<-env_lat[,4]
grid[45,1,]<-env_lat[,5]
grid[55,1,]<-env_lat[,6]
grid[65,1,]<-env_lat[,7]
grid[75,1,]<-env_lat[,8]
grid[85,1,]<-env_lat[,9]
image(grid[,1,],col=c("blue","green","red"))

```

we are printing this picture in a file 
```{r}
im<-array(NA,c(100,1,566))
im[5,1,]<-env_lat[,1]
im[6,1,]<-env_lat[,1]
im[15,1,]<-env_lat[,2]
im[16,1,]<-env_lat[,2]
im[25,1,]<-env_lat[,3]
im[26,1,]<-env_lat[,3]
im[35,1,]<-env_lat[,4]
im[36,1,]<-env_lat[,4]
im[45,1,]<-env_lat[,5]
im[46,1,]<-env_lat[,5]
im[55,1,]<-env_lat[,6]
im[56,1,]<-env_lat[,6]
im[65,1,]<-env_lat[,7]
im[66,1,]<-env_lat[,7]
im[75,1,]<-env_lat[,8]
im[76,1,]<-env_lat[,8]
im[85,1,]<-env_lat[,9]
im[86,1,]<-env_lat[,9]
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
png(file="logsutah.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```

we also need to put the proportions on a grid
```{r}
gridprop1<-gridprop2<-array(NA,c(100,2,566))
gridprop1[5,1,]<-prop1mat[,1]
gridprop1[15,1,]<-prop1mat[,2]
gridprop1[25,1,]<-prop1mat[,3]
gridprop1[35,1,]<-prop1mat[,4]
gridprop1[45,2,]<-prop1mat[,5]
gridprop1[55,1,]<-prop1mat[,6]
gridprop1[65,1,]<-prop1mat[,7]
gridprop1[75,1,]<-prop1mat[,8]
gridprop1[99,1,]<-prop1mat[,9]
#gridprop1[is.na(grid)]=NA
gridprop2[5,1,]<-prop2mat[,1]
gridprop2[15,1,]<-prop2mat[,2]
gridprop2[25,1,]<-prop2mat[,3]
gridprop2[35,1,]<-prop2mat[,4]
gridprop2[45,2,]<-prop2mat[,5]
gridprop2[55,1,]<-prop2mat[,6]
gridprop2[65,1,]<-prop2mat[,7]
gridprop2[75,1,]<-prop2mat[,8]
gridprop2[99,1,]<-prop2mat[,9]
#gridprop2[is.na(grid)]=NA
gridprop3=1-gridprop1-gridprop2
image(gridprop1[,1,])
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(10)[c(1, 10, 20)], xpd = NA)
image(gridprop2[,1,])
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(10)[c(1, 10, 20)], xpd = NA)
image(gridprop3[,1,])
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(10)[c(1, 10, 20)], xpd = NA)
```
okay, now we need to do a kriging to interpolate these proportions
```{r}
source("functions.R")
```
we need first to untrend the global proportion of each facies to perform the kriging
```{r}
gridprop1=gridprop1-p1
gridprop2=gridprop2-p2
```

```{r}
n=dim(gridprop1)
gridprop1krig<-prop_kriging(gridprop1,500,20,20,0.1)
gridprop2krig<-prop_kriging(gridprop2,500,20,20,0.1)
#now we need to retrend the results
gridprop1krig=gridprop1krig+p1
gridprop2krig=gridprop2krig+p2
err=1-gridprop1krig-gridprop2krig
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      if (err[x,y,z]<0){
        gridprop1krig[x,y,z]=gridprop1krig[x,y,z]+err[x,y,z]
      }
    }
  }
}

image(gridprop1krig[,1,],col=heat.colors(300))
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(300)[c(1, 200,300)], xpd = NA)
image(gridprop2krig[,1,],col=heat.colors(300))
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(300)[c(1, 200,300)], xpd = NA)
gridprop3krig<-1-gridprop2krig-gridprop1krig
image(gridprop3krig[,1,],col=heat.colors(300))
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("0",".5","1"), fill = heat.colors(300)[c(1 , 200,300)], xpd = NA)
```
test
```{r}
gridprop1krig[gridprop1krig>1]
gridprop2krig[gridprop2krig>1]
gridprop1krig[gridprop1krig<0]
gridprop2krig[gridprop2krig<0]
```

```{r}
library(mvtnorm)
```
we also need to correct when proportions are higher than one
```{r}
gridprop1krig[gridprop1krig>1]=1
gridprop2krig[gridprop2krig>1]=1
gridprop1krig[gridprop1krig<0]=0
gridprop2krig[gridprop2krig<0]=0
```

Ok now we can compute the corresponding thresholds for the simulation

we have no correlation between the Gaussian random functions, we have thus a simplification and can directly compute q2, with no approximation
```{r}
n=dim(gridprop1)
q1tottest=q2tottest=array(NA,c(100,2,566))
Sigma1 = matrix(c(1,rho,rho,1),2,2)
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      q1tottest[x,y,z]=qnorm(gridprop1krig[x,y,z])
      if (gridprop3krig[x,y,z]!=0) {
      q2tottest[x,y,z]=qnorm(1-(gridprop2krig[x,y,z]/(1-gridprop1krig[x,y,z])))
      }
      else{q2tottest[x,y,z]=-Inf}
    }
  }
}
image(q1tottest[,1,],col=heat.colors(300))
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("-Inf","0","Inf"), fill = heat.colors(300)[c(1, 150,300)], xpd = NA)
image(q2tottest[,1,],col=heat.colors(300))
legend(grconvertX(0.5, "device"), grconvertY(1, "device"), 
     c("-Inf","0","Inf"), fill = heat.colors(300)[c(1, 150,300)], xpd = NA)
```
we can check that the threshods on the final grid are equivalent to the one originally computed on the logs
```{r}
q1[1,1]
q1tottest[5,1,1]
q2[1,1]
q2tottest[5,1,1]
```

Once the thresholds are computed, we can do the simulation (unconditional for now)

```{r}
rv1=5
rv2=5
rh=800
b1=0
b2=0
rho=0
Y1<-sim3d(rv1,rh,rh,b1,100,0.1,566,20,100,20,2)
Y2<-sim3d(rv2,rh,rh,b2,100,0.1,566,20,100,20,2)
fac<-Y1
fac[]=3
fac[Y1<q1tottest]=1
fac[Y1>q1tottest & Y2>q2tottest]=2
image(grid[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,2,],col=c("darkslategray4","burlywood2","brown3"))
```
ok let's now do a conditional simulation
for the gibbs sampling we dont need the kriged proportions

```{r}
env_lat[is.na(env_lat)]=Inf # with 0, the gibbs sampler will find random values
grid<-array(NA,c(100,2,566))
grid[5,1,]<-env_lat[,1]
grid[15,1,]<-env_lat[,2]
grid[25,1,]<-env_lat[,3]
grid[35,1,]<-env_lat[,4]
grid[45,2,]<-env_lat[,5]
grid[55,1,]<-env_lat[,6]
grid[65,1,]<-env_lat[,7]
grid[75,1,]<-env_lat[,8]
grid[99,1,]<-env_lat[,9]
i1grid<-grid
i1grid[grid==2]=0
i1grid[grid==3]=0
image(i1grid[,1,])
q1map<-array(NA,c(100,2,566))
q1map[5,1,]<-q1[,1]
q1map[15,1,]<-q1[,2]
q1map[25,1,]<-q1[,3]
q1map[35,1,]<-q1[,4]
q1map[45,2,]<-q1[,5]
q1map[55,1,]<-q1[,6]
q1map[65,1,]<-q1[,7]
q1map[75,1,]<-q1[,8]
q1map[99,1,]<-q1[,9]

Y1gibbs<-gibbs_sampling_map(i1grid,q1map,0.1,20,10,rv1,rh,rh,b1,3)
image(Y1gibbs[,1,])

image(i1grid[,1,])
image(Y1gibbs[,1,])
image(q1map[,1,])
```
test
```{r}
test<-gibbs_sampling_map(i1grid,q1map,0.1,20,10,rv1,rh,rh,b1,1)
image(test[,1,],col=heat.colors(300))
```

```{r}
i3grid<-grid
i3grid[grid==1]=Inf
i3grid[grid==3]=1
i3grid[grid==2]=0
q2map<-array(NA,c(100,2,566))
q2map[5,1,]<-q2tottest[5,1,]
q2map[15,1,]<-q2tottest[15,1,]
q2map[25,1,]<-q2tottest[25,1,]
q2map[35,1,]<-q2tottest[35,1,]
q2map[45,2,]<-q2tottest[45,2,]
q2map[55,1,]<-q2tottest[55,1,]
q2map[65,1,]<-q2tottest[65,1,]
q2map[75,1,]<-q2tottest[75,1,]
q2map[99,1,]<-q2tottest[99,1,]

Y2gibbs<-gibbs_sampling_map(i3grid,q2map,0.1,20,10,rv2,rh,rh,b2,3)
image(Y2gibbs[,1,])
test<-Y2gibbs
test[Y2gibbs<q2map]=1
test[Y2gibbs>q2map]=0
image(test[,1,])
image(i3grid[,1,])
image(Y2gibbs[,1,])
```

now that we have the gibbs sampling, we need to krig the difference with the unconditional simulation

```{r}
E1=Y1gibbs-Y1
E2=Y2gibbs-Y2
E1k<-prop_kriging(E1,rh,20,20,0.1)
E2k<-prop_kriging(E2,rh,20,20,0.1)
Y1c=Y1+E1k
Y2c=Y2+E2k
facc<-Y1c
facc[]=3
facc[Y1c<q1tottest]=1
facc[Y1c>q1tottest & Y2c>q2tottest]=2
image(facc[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(grid[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(E1k[,1,],col=heat.colors(300))
image(E2k[,1,],col=heat.colors(300))
```

ok now we can look at the diagenesis, we do it separately because in this example,
the Gaussian function representing the diagenesis is not correlated to the Gaussian functions of the depositional facies
```{r}
diag_lat<-data.matrix(read.table("diagclastic.txt"))
diag_lat[diag_lat==0]=3
d1=d2=d3=diag_lat
d1[diag_lat!=1]=0
d2[diag_lat==2]=1
d2[diag_lat!=2]=0
d3[diag_lat==3]=1
d3[diag_lat!=3]=0
pd1=mean(d1,na.rm=TRUE)
pd2=mean(d2,na.rm=TRUE)
pd3=mean(d3,na.rm=TRUE)
c(pd1,pd2,pd3)
image(t(diag_lat),col=c("blue","green","red"))
```

load the diagenesis
```{r}

im<-array(NA,c(100,1,566))
im[5,1,]<-env_lat[,1]
im[6,1,]<-env_lat[,1]
im[8,1,]<-diag_lat[,1]+3
im[9,1,]<-diag_lat[,1]+3
im[15,1,]<-env_lat[,2]
im[16,1,]<-env_lat[,2]
im[18,1,]<-diag_lat[,2]+3
im[19,1,]<-diag_lat[,2]+3
im[25,1,]<-env_lat[,3]
im[26,1,]<-env_lat[,3]
im[28,1,]<-diag_lat[,3]+3
im[29,1,]<-diag_lat[,3]+3
im[35,1,]<-env_lat[,4]
im[36,1,]<-env_lat[,4]
im[38,1,]<-diag_lat[,4]+3
im[39,1,]<-diag_lat[,4]+3
im[45,1,]<-env_lat[,5]
im[46,1,]<-env_lat[,5]
im[48,1,]<-diag_lat[,5]+3
im[49,1,]<-diag_lat[,5]+3
im[55,1,]<-env_lat[,6]
im[56,1,]<-env_lat[,6]
im[58,1,]<-diag_lat[,6]+3
im[59,1,]<-diag_lat[,6]+3
im[65,1,]<-env_lat[,7]
im[66,1,]<-env_lat[,7]
im[68,1,]<-diag_lat[,7]+3
im[69,1,]<-diag_lat[,7]+3
im[75,1,]<-env_lat[,8]
im[76,1,]<-env_lat[,8]
im[78,1,]<-diag_lat[,8]+3
im[79,1,]<-diag_lat[,8]+3
im[85,1,]<-env_lat[,9]
im[86,1,]<-env_lat[,9]
im[88,1,]<-diag_lat[,9]+3
im[89,1,]<-diag_lat[,9]+3
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow","black"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
png(file="logsutah.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow","black"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```

simulation now with diagenesis! 
we first need to compute the relative proportions
```{r}
pd1e1=mean(i1*d1,na.rm=TRUE)/p1
pd2e1=mean(i1*d2,na.rm=TRUE)/p1
pd3e1=mean(i1*d3,na.rm=TRUE)/p1
pd1e2=mean(i2*d1,na.rm=TRUE)/p2
pd2e2=mean(i2*d2,na.rm=TRUE)/p2
pd3e2=mean(i2*d3,na.rm=TRUE)/p2
pd1e3=mean(i3*d1,na.rm=TRUE)/p3
pd2e3=mean(i3*d2,na.rm=TRUE)/p3
pd3e3=mean(i3*d3,na.rm=TRUE)/p3
matrix(c(pd1e1,pd2e1,pd3e1,pd1e2,pd2e2,pd3e2,pd1e3,pd2e3,pd3e3),3,3)
```

now we need to find the thresholds
given that d1 is only found in e2, its threshold is computed with its general proportion
but this is not the case for d2
we need first to compute the thresholds of d2 because will impact threshold of d1
```{r}
qd2e2=qnorm(1-pd2e2)
qd1e2=qnorm(1-pd2e2-pd1e2)
qd2e3=qnorm(1-pd2e3)
c(qd2e2,qd1e2,qd2e3)
```
Now I can do the simulation, with in mind that I also need to condition it! 
but it should be simpler than previously. 
I still need to rewrite the gibbs sampling for that new truncation rule
```{r}
dgrid<-array(NA,c(100,2,566))
diag_lat[is.na(diag_lat)]=Inf
dgrid[5,1,]=diag_lat[,1]
dgrid[15,1,]<-diag_lat[,2]
dgrid[25,1,]<-diag_lat[,3]
dgrid[35,1,]<-diag_lat[,4]
dgrid[45,2,]<-diag_lat[,5]
dgrid[55,1,]<-diag_lat[,6]
dgrid[65,1,]<-diag_lat[,7]
dgrid[75,1,]<-diag_lat[,8]
dgrid[99,1,]<-diag_lat[,9]
Y3gibbs<-gibbs_sampling_diag2(grid,dgrid,c(qd1e2,qd2e2,qd2e3),0.1,20,10,rv2,rh,rh,b2,3)
image(Y3gibbs[,1,])
facdg<-Y3gibbs
facdg[!is.na(facdg)]=3
facdg[ grid==2 & Y3gibbs<qd2e2 & Y3gibbs>qd1e2]=1
facdg[ grid==2 & Y3gibbs>qd2e2 ]=2
facdg[ grid==3 & Y3gibbs>qd2e3 ]=2
image(facdg[,1,],col=c("green","yellow","black"))
image(dgrid[,1,],col=c("green","yellow","black"))
```

```{r}
Y3<-sim3d(rv2,rh,rh,b2,100,0.1,566,20,100,20,2)
```
```{r}
E3=Y3gibbs-Y3
E3k<-prop_kriging(E3,rh,20,20,0.1)
Y3c=Y3+E3k
facd<-Y3c
facd[]=3
facd[ facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=1
facd[ facc==2 & Y3c>qd2e2 ]=2
facd[ facc==3 & Y3c>qd2e3 ]=2
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow","black"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
image(facd[,1,],col=c("green","yellow","black"))
fac_comb=facc
fac_comb[facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=4
fac_comb[facc==2 & Y3c>qd2e2 ]=5
fac_comb[facc==3 & Y3c>qd2e3 ]=5
image(facc[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac_comb[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow"))
```
Now that it is working, we can make a beautiful simulation in three dimensions
```{r}
par(bg="transparent")
n=dim(gridprop1)
gridprop1<-gridprop2<-err<-array(NA,c(100,20,566))
gridprop1[5,10,]<-prop1mat[,1]
gridprop1[15,9,]<-prop1mat[,2]
gridprop1[25,10,]<-prop1mat[,3]
gridprop1[35,9,]<-prop1mat[,4]
gridprop1[45,10,]<-prop1mat[,5]
gridprop1[55,9,]<-prop1mat[,6]
gridprop1[65,10,]<-prop1mat[,7]
gridprop1[75,9,]<-prop1mat[,8]
gridprop1[99,10,]<-prop1mat[,9]
#gridprop1[is.na(grid)]=NA
gridprop2[5,10,]<-prop2mat[,1]
gridprop2[15,9,]<-prop2mat[,2]
gridprop2[25,10,]<-prop2mat[,3]
gridprop2[35,9,]<-prop2mat[,4]
gridprop2[45,10,]<-prop2mat[,5]
gridprop2[55,9,]<-prop2mat[,6]
gridprop2[65,10,]<-prop2mat[,7]
gridprop2[75,9,]<-prop2mat[,8]
gridprop2[99,10,]<-prop2mat[,9]
#gridprop2[is.na(grid)]=NA
gridprop3=1-gridprop1-gridprop2
gridprop1=gridprop1-p1
gridprop2=gridprop2-p2
gridprop1krig<-prop_kriging(gridprop1,500,20,20,0.1)
gridprop2krig<-prop_kriging(gridprop2,500,20,20,0.1)
# we are trying here to have the same proportions on every y plane (because no information on them but still want non sationarity)
for (y in 1:20){
  if (y!=10 & y!=9){
    gridprop1krig[,y,]=gridprop1krig[,10,]
    gridprop2krig[,y,]=gridprop2krig[,10,]
  }
}

#now we need to retrend the results
gridprop1krig=gridprop1krig+p1
gridprop2krig=gridprop2krig+p2
gridprop3krig=1-gridprop1krig-gridprop2krig # there is a problem  here to solve
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      err[x,y,z]=gridprop1krig[x,y,z]+gridprop2krig[x,y,z]
      if (err[x,y,z]>1){
        gridprop1krig[x,y,z]=gridprop1krig[x,y,z]+(1-err[x,y,z])
        gridprop3krig[x,y,z]=0
      }
    }
  }
}

n=dim(gridprop1)
q1tottest=q2tottest=array(NA,c(100,20,566))
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      q1tottest[x,y,z]=qnorm(gridprop1krig[x,y,z])
      if (gridprop3krig[x,y,z]>0) {
        q2tottest[x,y,z]=qnorm(1-(gridprop2krig[x,y,z]/(1-gridprop1krig[x,y,z])))
      }
      else{q2tottest[x,y,z]=-Inf}
    }
  }
}

rv1=5
rv2=5
rv3=2
rh=300
rh3=150
b1=0
b2=0
rho=0
Y1<-sim3d(rv1,rh,rh,b1,100,0.1,566,20,100,20,20)
Y2<-sim3d(rv2,rh,rh,b2,100,0.1,566,20,100,20,20)
Y3<-sim3d(rv3,rh3,rh3,b2,100,0.1,566,20,100,20,20)
fac<-Y1
fac[]=3
fac[Y1<q1tottest]=1
fac[Y1>q1tottest & Y2>q2tottest]=2
image(grid[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,2,],col=c("darkslategray4","burlywood2","brown3"))
env_lat[is.na(env_lat)]=Inf # with 0, the gibbs sampler will find random values
grid<-array(NA,c(100,20,566))
grid[5,10,]<-env_lat[,1]
grid[15,9,]<-env_lat[,2]
grid[25,10,]<-env_lat[,3]
grid[35,9,]<-env_lat[,4]
grid[45,10,]<-env_lat[,5]
grid[55,9,]<-env_lat[,6]
grid[65,10,]<-env_lat[,7]
grid[75,9,]<-env_lat[,8]
grid[99,10,]<-env_lat[,9]
i1grid<-grid
i1grid[grid==2]=0
i1grid[grid==3]=0
image(i1grid[,1,])
q1map<-array(NA,c(100,20,566))
q1map[5,10,]<-q1tottest[5,10,]
q1map[15,9,]<-q1tottest[15,9,]
q1map[25,10,]<-q1tottest[25,10,]
q1map[35,9,]<-q1tottest[35,9,]
q1map[45,10,]<-q1tottest[45,10,]
q1map[55,9,]<-q1tottest[55,9,]
q1map[65,10,]<-q1tottest[65,10,]
q1map[75,9,]<-q1tottest[75,9,]
q1map[99,10,]<-q1tottest[99,10,]

Y1gibbs<-gibbs_sampling_map(i1grid,q1map,0.1,20,10,rv1,rh,rh,b1,3)
image(Y1gibbs[,10,])
image(i1grid[,10,])
image(Y1gibbs[,10,])
image(q1map[,10,])

i3grid<-grid
i3grid[grid==1]=Inf
i3grid[grid==3]=1
i3grid[grid==2]=0
q2map<-array(NA,c(100,20,566))
q2map[5,10,]<-q2tottest[5,10,]
q2map[15,9,]<-q2tottest[15,9,]
q2map[25,10,]<-q2tottest[25,10,]
q2map[35,9,]<-q2tottest[35,9,]
q2map[45,10,]<-q2tottest[45,10,]
q2map[55,9,]<-q2tottest[55,9,]
q2map[65,10,]<-q2tottest[65,10,]
q2map[75,9,]<-q2tottest[75,9,]
q2map[99,10,]<-q2tottest[99,10,]

Y2gibbs<-gibbs_sampling_map(i3grid,q2map,0.1,20,10,rv2,rh,rh,b2,3)
image(Y2gibbs[,10,])

E1=Y1gibbs-Y1
E2=Y2gibbs-Y2
E1k<-prop_kriging(E1,rh,20,20,0.1)
E2k<-prop_kriging(E2,rh,20,20,0.1)
Y1c=Y1+E1k
Y2c=Y2+E2k
facc<-Y1c
facc[]=3
facc[Y1c<q1tottest]=1
facc[Y1c>q1tottest & Y2c>q2tottest]=2
image(facc[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(grid[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(E1k[,10,],col=heat.colors(300))
image(E2k[,10,],col=heat.colors(300))
dgrid<-array(NA,c(100,20,566))
diag_lat[is.na(diag_lat)]=Inf
dgrid[5,10,]=diag_lat[,1]
dgrid[15,9,]<-diag_lat[,2]
dgrid[25,10,]<-diag_lat[,3]
dgrid[35,9,]<-diag_lat[,4]
dgrid[45,10,]<-diag_lat[,5]
dgrid[55,9,]<-diag_lat[,6]
dgrid[65,10,]<-diag_lat[,7]
dgrid[75,9,]<-diag_lat[,8]
dgrid[99,10,]<-diag_lat[,9]
Y3gibbs<-gibbs_sampling_diag(grid,dgrid,c(qd1e2,qd2e2,qd2e3),0.1,20,10,rv2,rh,rh,b2,3)
image(Y3gibbs[,10,])
facdg<-Y3gibbs
facdg[!is.na(facdg)]=3
facdg[ grid==2 & Y3gibbs<qd2e2 & Y3gibbs>qd1e2]=1
facdg[ grid==2 & Y3gibbs>qd2e2 ]=2
facdg[ grid==3 & Y3gibbs>qd2e3 ]=2
image(facdg[,10,],col=c("green","yellow","black"))
image(dgrid[,10,],col=c("green","yellow","black"))
E3=Y3gibbs-Y3
E3k<-prop_kriging(E3,rh3,20,20,0.1)
Y3c=Y3+E3k
facd<-Y3c
facd[]=3
facd[ facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=1
facd[ facc==2 & Y3c>qd2e2 ]=2
facd[ facc==3 & Y3c>qd2e3 ]=2

image(facd[,10,],col=c("green","yellow","black"))
fac_comb=facc
fac_comb[facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=4
fac_comb[facc==2 & Y3c>qd2e2 ]=5
fac_comb[facc==3 & Y3c>qd2e3 ]=5
image(facc[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(fac_comb[,10,],col=c("darkslategray4","burlywood2","brown3","green","yellow"))
image(dgrid[,10,],col=c("green","yellow","black"))
```

```{r}
for ( z in 1:566){
  for ( y in 1:20){
    for ( x in 1:100){
      write(toString(c(fac_comb[x,y,z])),"utahsimu5.csv",append=TRUE)
    }
  }
}
```
we also want to export the data
```{r}
grid[grid==Inf]=NA
dgrid[dgrid==Inf]=NA
comb_grid<-grid
comb_grid[dgrid==1]=4
comb_grid[dgrid==2]=5
for ( z in 1:566){
  for ( y in 1:20){
    for ( x in 1:100){
      write(toString(c(comb_grid[x,y,z])),"utahdata.csv",append=TRUE)
    }
  }
}
```
we can also compute the resulting embedded transitions on those simulations
```{r}
i1real<-i2real<-i3real<-facc
i1real[facc!=1]=0
i2real[facc==2]=1
i2real[facc!=2]=0
i3real[facc==3]=1
i3real[facc!=3]=0
r12=embedded3d(i1real,i2real)
r13=embedded3d(i1real,i3real)
r21=embedded3d(i2real,i1real)
r23=embedded3d(i2real,i3real)
r31=embedded3d(i3real,i1real)
r32=embedded3d(i3real,i2real)
matrix(c(0,r12,r13,r21,0,r23,r31,r32,0),3,3,byrow=TRUE)
r12d=embedded2d(i1,i2)
r13d=embedded2d(i1,i3)
r21d=embedded2d(i2,i1)
r23d=embedded2d(i2,i3)
r31d=embedded2d(i3,i1)
r32d=embedded2d(i3,i2)
matrix(c(0,r12d,r13d,r21d,0,r23d,r31d,r32d,0),3,3,byrow=TRUE)
```

```{r}
real3<-matrix(c(0,r12,r13,r21,0,r23,r31,r32,0),3,3,byrow=TRUE)
```
```{r}
(real1+real2+real3)/3
```