### 2018-10-16 main_bookcliff.Rmd
cleaning

---
title: "R Notebook"
output: html_notebook
---
facies1=distal lower shoreface heteroliths & offshore mudstones
facies2=proximal lower shoreface sandstones
facies3=foreshore & upper shoreface sandstones

This case study is not stationary so we are importing functions
redefined in a non stationary case
```{r}
source("nonstation_functions.R")
```
```{r}
library(mvtnorm)
```
importing depositional facies
```{r}
i1=i2=i3=env_lat<-data.matrix(read.table("envclastic.txt"))
i1[env_lat!=1]=0
i2[env_lat==2]=1
i2[env_lat!=2]=0
i3[env_lat==3]=1
i3[env_lat!=3]=0
p1=mean(i1,na.rm=TRUE)
p2=mean(i2,na.rm=TRUE)
p3=mean(i3,na.rm=TRUE)
c(p1,p2,p3)
image(t(env_lat),col=c("blue","green","red"))
```

computing transiograms on the data
```{r}
par(mfrow=c(3,3))
t11exp=tr_2d_vert(i1,i1,200,566,56.6)
t12exp=tr_2d_vert(i1,i2,200,566,56.6)
t13exp=tr_2d_vert(i1,i3,200,566,56.6)
t21exp=tr_2d_vert(i2,i1,200,566,56.6)
t22exp=tr_2d_vert(i2,i2,200,566,56.6)
t23exp=tr_2d_vert(i2,i3,200,566,56.6)
t31exp=tr_2d_vert(i3,i1,200,566,56.6)
t32exp=tr_2d_vert(i3,i2,200,566,56.6)
t33exp=tr_2d_vert(i3,i3,200,566,56.6)
```

Because of the non stationarity, it is more relevant to computed the transition rates between facies rather than the
transiograms
we can thus fit the method with the transition rates

Here are thresholds for the depositional facies according to the truncation rule
```{r}
q1g=qnorm(p1)
q2g=qnorm(1-(p2/(1-p1)))
```

covariance model for the Gaussian functions
```{r}
model<-function(h,range,b){return(exp(-abs(h/range)^2)*cos(b*h) )}
```

Model embedded transition probabilities
```{r}
rv1=5
rv2=5
rh=800
b1=0
b2=0
rho=0
rv1=0.6
rv2=0.7
Sigma = diag(1,4)
d = 1/10
Sigma[1,2]=Sigma[2,1]= Sigma[3,4]=Sigma[4,3]=rho
Sigma[1,3]=Sigma[3,1]=model(d,rv1,b1)
Sigma[1,4]=Sigma[4,1]=(rho)*model(abs(d),rv1,b1)
Sigma[2,3]=Sigma[3,2]=(rho)*model(abs(d),rv1,b1)
Sigma[2,4]=Sigma[4,2]=(rho^2)*model(d,rv1,b1)+(1-(rho^2))*model(d,rv2,b2)
r11_model=(1-pmvnorm(lower=c(-Inf,-Inf,-Inf,-Inf),upper=c(q1g,Inf,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p1)
r12_model=pmvnorm(lower=c(-Inf,-Inf,q1g,q2g),upper=c(q1g,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/(p1)
r13_model=pmvnorm(lower=c(-Inf,-Inf,q1g,-Inf),upper=c(q1g,Inf,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p1
r21_model=pmvnorm(lower=c(q1g,q2g,-Inf,-Inf),upper=c(Inf,Inf,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p2
r22_model=(1-pmvnorm(lower=c(q1g,q2g,q1g,q2g),upper=c(Inf,Inf,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/p2)
r23_model=pmvnorm(lower=c(q1g,q2g,q1g,-Inf),upper=c(Inf,Inf,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p2
r31_model=pmvnorm(lower=c(q1g,-Inf,-Inf,-Inf),upper=c(Inf,q2g,q1g,Inf),mean=c(0,0,0,0),sigma=Sigma)/p3
r32_model=pmvnorm(lower=c(q1g,-Inf,q1g,q2g),upper=c(Inf,q2g,Inf,Inf),mean=c(0,0,0,0),sigma=Sigma)/p3
r33_model=(1-pmvnorm(lower=c(q1g,-Inf,q1g,-Inf),upper=c(Inf,q2g,Inf,q2g),mean=c(0,0,0,0),sigma=Sigma)/p3)
matrix(c(0,r12_model/r11_model,r13_model/r11_model,r21_model/r22_model,0,r23_model/r22_model,r31_model/r33_model,r32_model/r33_model,0),3,3,byrow=TRUE)

```
The resulting facies thicknesses are

```{r}
(1/r11_model)*0.1
(1/r22_model)*0.1
(1/r33_model)*0.1
```
and the experimental embedded transitions are

```{r}
r11exp=-(1-t11exp[1])/0.1
r12exp=t12exp[1]/0.1
r13exp=t13exp[1]/0.1
r21exp=t21exp[1]/0.1
r22exp=-(1-t22exp[1])/0.1
r23exp=t23exp[1]/0.1
r31exp=t31exp[1]/0.1
r32exp=t32exp[1]/0.1
r33exp=-(1-t33exp[1])/0.1
matrix(c(r11exp,r12exp,r13exp,r21exp,r22exp,r23exp,r31exp,r32exp,r33exp),3,3)
```

and mean thicknesses
```{r}
-(1/r11exp)*0.1
-(1/r22exp)*0.1
-(1/r33exp)*0.1
```
Once the model is fitted, we can do the simulation,
But we have seen that the proportions are not stationary and we need to take that into account in the simulation

therefore, we need to compute the proportion on each horizontal 
```{r}
prop1=prop2=prop3=numeric(nrow(env_lat))
for (i in 1:nrow(env_lat)){
  prop1[i]=mean(i1[i,],na.rm=TRUE)
  prop2[i]=mean(i2[i,],na.rm=TRUE)
  prop3[i]=mean(i3[i,],na.rm=TRUE)
}
plot(1:nrow(i1),prop1,type = "l")
plot(1:nrow(i1),prop2,type = "l")
plot(1:nrow(i1),prop3,type = "l")
```
ok we are now smoothing the curves with moving average
```{r}
f50=rep(1/50,50)
prop1filter<-filter(prop1,f50,sides=2)
prop1filter[542:566]=prop1[542:566]
prop1filter[1:24]=prop1[1:24]
plot(1:nrow(i1),prop1filter,type = "l")

prop3filter<-filter(prop3,f50,sides=2)
prop3filter[542:566]=prop3[542:566]
prop3filter[1:24]=prop3[1:24]
plot(1:nrow(i1),prop3filter,type = "l")
prop2filter=1-prop1filter-prop3filter
plot(1:nrow(i1),prop2filter,type = "l")

```
we want to plot an image with those curves
```{r}
png(file="PCvert.png",width=2000,height = 800,res=300,antialias = )
par(mar=c(0,0,0,0))
n=nrow(i1)
plot(1:n,prop2filter+prop3filter,type="l",axes=FALSE,ann=FALSE)
lines(1:n,prop3filter)
lines(1:n,rep(1,n))
lines(1:n,rep(0,n))
```

therefore, we need to compute the proportion on each vertical 
```{r}
prop1lat=prop2lat=prop3lat=numeric(ncol(env_lat))
for (i in 1:ncol(env_lat)){
  prop1lat[i]=mean(i1[,i],na.rm=TRUE)
  prop2lat[i]=mean(i2[,i],na.rm=TRUE)
  prop3lat[i]=mean(i3[,i],na.rm=TRUE)
}
plot(1:ncol(i1),prop1lat,type = "p")
plot(1:ncol(i1),prop2lat,type = "p")
plot(1:ncol(i1),prop3lat,type = "p")
```
now doing a good figure
```{r}
png(file="PChor.png",width=2000,height = 800,res=300,antialias = )
par(mar=c(0,0,0,0))
n=ncol(i1)
plot(1:n,prop2lat+prop3lat,type="p",axes=FALSE,ann=FALSE,ylim=c(0,1))
lines(1:n,prop3lat,type="p")
lines(1:n,rep(1,n))
lines(1:n,rep(0,n))
```

we are now doing the mean of the lateral and vertical proportion curves
```{r}
prop1mat=prop2mat=prop3mat=matrix(0,nrow=nrow(env_lat),ncol=ncol(env_lat))
for ( i in 1:nrow(env_lat)){
  for (j in 1:ncol(env_lat)){
    prop1mat[i,j]=(prop1filter[i]+prop1lat[j])/2
    prop2mat[i,j]=(prop2filter[i]+prop2lat[j])/2
    #prop3mat[i,j]=(prop3filter[i]+prop3lat[j])/2
  }
}
prop3mat=1-prop2mat-prop1mat
prop3mat[prop3mat<0]=0
image(prop1mat)
image(prop2mat)
image(prop3mat)
```
Once the local proportions have been computed on each data point, we can interpolate those proportions in the grid

we create a grid (3d) to simulate the depositional facies

```{r}
grid<-array(NA,c(100,1,566))
grid[5,1,]<-env_lat[,1]
grid[15,1,]<-env_lat[,2]
grid[25,1,]<-env_lat[,3]
grid[35,1,]<-env_lat[,4]
grid[45,1,]<-env_lat[,5]
grid[55,1,]<-env_lat[,6]
grid[65,1,]<-env_lat[,7]
grid[75,1,]<-env_lat[,8]
grid[85,1,]<-env_lat[,9]
image(grid[,1,],col=c("blue","green","red"))

```

we are printing this picture in a file 
```{r}
im<-array(NA,c(100,1,566))
im[5,1,]<-env_lat[,1]
im[6,1,]<-env_lat[,1]
im[15,1,]<-env_lat[,2]
im[16,1,]<-env_lat[,2]
im[25,1,]<-env_lat[,3]
im[26,1,]<-env_lat[,3]
im[35,1,]<-env_lat[,4]
im[36,1,]<-env_lat[,4]
im[45,1,]<-env_lat[,5]
im[46,1,]<-env_lat[,5]
im[55,1,]<-env_lat[,6]
im[56,1,]<-env_lat[,6]
im[65,1,]<-env_lat[,7]
im[66,1,]<-env_lat[,7]
im[75,1,]<-env_lat[,8]
im[76,1,]<-env_lat[,8]
im[85,1,]<-env_lat[,9]
im[86,1,]<-env_lat[,9]
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
png(file="logsutah.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```

and the diagenesis
```{r}
diag_lat<-data.matrix(read.table("diagclastic.txt"))
diag_lat[diag_lat==0]=3
d1=d2=d3=diag_lat
d1[diag_lat!=1]=0
d2[diag_lat==2]=1
d2[diag_lat!=2]=0
d3[diag_lat==3]=1
d3[diag_lat!=3]=0
pd1=mean(d1,na.rm=TRUE)
pd2=mean(d2,na.rm=TRUE)
pd3=mean(d3,na.rm=TRUE)
c(pd1,pd2,pd3)
image(t(diag_lat),col=c("blue","green","red"))
```

load the diagenesis data 
```{r}

im<-array(NA,c(100,1,566))
im[5,1,]<-env_lat[,1]
im[6,1,]<-env_lat[,1]
im[8,1,]<-diag_lat[,1]+3
im[9,1,]<-diag_lat[,1]+3
im[15,1,]<-env_lat[,2]
im[16,1,]<-env_lat[,2]
im[18,1,]<-diag_lat[,2]+3
im[19,1,]<-diag_lat[,2]+3
im[25,1,]<-env_lat[,3]
im[26,1,]<-env_lat[,3]
im[28,1,]<-diag_lat[,3]+3
im[29,1,]<-diag_lat[,3]+3
im[35,1,]<-env_lat[,4]
im[36,1,]<-env_lat[,4]
im[38,1,]<-diag_lat[,4]+3
im[39,1,]<-diag_lat[,4]+3
im[45,1,]<-env_lat[,5]
im[46,1,]<-env_lat[,5]
im[48,1,]<-diag_lat[,5]+3
im[49,1,]<-diag_lat[,5]+3
im[55,1,]<-env_lat[,6]
im[56,1,]<-env_lat[,6]
im[58,1,]<-diag_lat[,6]+3
im[59,1,]<-diag_lat[,6]+3
im[65,1,]<-env_lat[,7]
im[66,1,]<-env_lat[,7]
im[68,1,]<-diag_lat[,7]+3
im[69,1,]<-diag_lat[,7]+3
im[75,1,]<-env_lat[,8]
im[76,1,]<-env_lat[,8]
im[78,1,]<-diag_lat[,8]+3
im[79,1,]<-diag_lat[,8]+3
im[85,1,]<-env_lat[,9]
im[86,1,]<-env_lat[,9]
im[88,1,]<-diag_lat[,9]+3
im[89,1,]<-diag_lat[,9]+3
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow","black"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
png(file="logsutah.png",width=4000,height = 2500,res=300,antialias = )
image(x=(1:100)*200,y=(1:566)/10,z=im[,1,],col=c("darkslategray4","burlywood2","brown3","green","yellow","black"),xlab="Lateral distance (m)",ylab="Vertical distance (m)")
```
we suppose that the diagenetic facies have a non stationarity only resuting from the non stationarity of the depositional facies

we compute the relative proportions of diagenesis in depositional facies
```{r}
pd1e1=mean(i1*d1,na.rm=TRUE)/p1
pd2e1=mean(i1*d2,na.rm=TRUE)/p1
pd3e1=mean(i1*d3,na.rm=TRUE)/p1
pd1e2=mean(i2*d1,na.rm=TRUE)/p2
pd2e2=mean(i2*d2,na.rm=TRUE)/p2
pd3e2=mean(i2*d3,na.rm=TRUE)/p2
pd1e3=mean(i3*d1,na.rm=TRUE)/p3
pd2e3=mean(i3*d2,na.rm=TRUE)/p3
pd3e3=mean(i3*d3,na.rm=TRUE)/p3
matrix(c(pd1e1,pd2e1,pd3e1,pd1e2,pd2e2,pd3e2,pd1e3,pd2e3,pd3e3),3,3)
```

now we need to find the corresponding thresholds
given that d1 is only found in e2, its threshold is computed with its general proportion
but this is not the case for d2
we need first to compute the thresholds of d2 because will impact threshold of d1
```{r}
qd2e2=qnorm(1-pd2e2)
qd1e2=qnorm(1-pd2e2-pd1e2)
qd2e3=qnorm(1-pd2e3)
c(qd2e2,qd1e2,qd2e3)
```
Now I can do the simulation
The gibbs sampling is different that in the Latemar because the truncation rule is different and adptative

```{r}
par(bg="transparent")
n=dim(gridprop1)
gridprop1<-gridprop2<-err<-array(NA,c(100,20,566))
gridprop1[5,10,]<-prop1mat[,1]
gridprop1[15,9,]<-prop1mat[,2]
gridprop1[25,10,]<-prop1mat[,3]
gridprop1[35,9,]<-prop1mat[,4]
gridprop1[45,10,]<-prop1mat[,5]
gridprop1[55,9,]<-prop1mat[,6]
gridprop1[65,10,]<-prop1mat[,7]
gridprop1[75,9,]<-prop1mat[,8]
gridprop1[99,10,]<-prop1mat[,9]
#gridprop1[is.na(grid)]=NA
gridprop2[5,10,]<-prop2mat[,1]
gridprop2[15,9,]<-prop2mat[,2]
gridprop2[25,10,]<-prop2mat[,3]
gridprop2[35,9,]<-prop2mat[,4]
gridprop2[45,10,]<-prop2mat[,5]
gridprop2[55,9,]<-prop2mat[,6]
gridprop2[65,10,]<-prop2mat[,7]
gridprop2[75,9,]<-prop2mat[,8]
gridprop2[99,10,]<-prop2mat[,9]
#gridprop2[is.na(grid)]=NA
gridprop3=1-gridprop1-gridprop2
gridprop1=gridprop1-p1
gridprop2=gridprop2-p2
gridprop1krig<-prop_kriging(gridprop1,500,20,20,0.1)
gridprop2krig<-prop_kriging(gridprop2,500,20,20,0.1)
# we are trying here to have the same proportions on every y plane (because no information on them but still want non sationarity)
for (y in 1:20){
  if (y!=10 & y!=9){
    gridprop1krig[,y,]=gridprop1krig[,10,]
    gridprop2krig[,y,]=gridprop2krig[,10,]
  }
}

#now we need to retrend the results
gridprop1krig=gridprop1krig+p1
gridprop2krig=gridprop2krig+p2
gridprop3krig=1-gridprop1krig-gridprop2krig # there is a problem  here to solve
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      err[x,y,z]=gridprop1krig[x,y,z]+gridprop2krig[x,y,z]
      if (err[x,y,z]>1){
        gridprop1krig[x,y,z]=gridprop1krig[x,y,z]+(1-err[x,y,z])
        gridprop3krig[x,y,z]=0
      }
    }
  }
}

# now that we have the proportions, we can compute the corresponding thresholds
# because there are no correlations between the gaussian functions, the thresholds can be computed analytically
n=dim(gridprop1)
q1tottest=q2tottest=array(NA,c(100,20,566))
for (x in 1:n[1]){
  for(y in 1:n[2]){
    for( z in 1:n[3]){
      q1tottest[x,y,z]=qnorm(gridprop1krig[x,y,z])
      if (gridprop3krig[x,y,z]>0) {
        q2tottest[x,y,z]=qnorm(1-(gridprop2krig[x,y,z]/(1-gridprop1krig[x,y,z])))
      }
      else{q2tottest[x,y,z]=-Inf}
    }
  }
}

# simulation
rv1=5
rv2=5
rv3=2
rh=300
rh3=150
b1=0
b2=0
rho=0
Y1<-sim3d(rv1,rh,rh,b1,100,0.1,566,20,100,20,20)
Y2<-sim3d(rv2,rh,rh,b2,100,0.1,566,20,100,20,20)
Y3<-sim3d(rv3,rh3,rh3,b2,100,0.1,566,20,100,20,20)
fac<-Y1
fac[]=3
fac[Y1<q1tottest]=1
fac[Y1>q1tottest & Y2>q2tottest]=2
image(grid[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,1,],col=c("darkslategray4","burlywood2","brown3"))
image(fac[,2,],col=c("darkslategray4","burlywood2","brown3"))
env_lat[is.na(env_lat)]=Inf # with 0, the gibbs sampler will find random values
grid<-array(NA,c(100,20,566))
grid[5,10,]<-env_lat[,1]
grid[15,9,]<-env_lat[,2]
grid[25,10,]<-env_lat[,3]
grid[35,9,]<-env_lat[,4]
grid[45,10,]<-env_lat[,5]
grid[55,9,]<-env_lat[,6]
grid[65,10,]<-env_lat[,7]
grid[75,9,]<-env_lat[,8]
grid[99,10,]<-env_lat[,9]
i1grid<-grid
i1grid[grid==2]=0
i1grid[grid==3]=0
image(i1grid[,1,])
q1map<-array(NA,c(100,20,566))
q1map[5,10,]<-q1tottest[5,10,]
q1map[15,9,]<-q1tottest[15,9,]
q1map[25,10,]<-q1tottest[25,10,]
q1map[35,9,]<-q1tottest[35,9,]
q1map[45,10,]<-q1tottest[45,10,]
q1map[55,9,]<-q1tottest[55,9,]
q1map[65,10,]<-q1tottest[65,10,]
q1map[75,9,]<-q1tottest[75,9,]
q1map[99,10,]<-q1tottest[99,10,]

Y1gibbs<-gibbs_sampling_map(i1grid,q1map,0.1,20,10,rv1,rh,rh,b1,3)
image(Y1gibbs[,10,])
image(i1grid[,10,])
image(Y1gibbs[,10,])
image(q1map[,10,])

i3grid<-grid
i3grid[grid==1]=Inf
i3grid[grid==3]=1
i3grid[grid==2]=0
q2map<-array(NA,c(100,20,566))
q2map[5,10,]<-q2tottest[5,10,]
q2map[15,9,]<-q2tottest[15,9,]
q2map[25,10,]<-q2tottest[25,10,]
q2map[35,9,]<-q2tottest[35,9,]
q2map[45,10,]<-q2tottest[45,10,]
q2map[55,9,]<-q2tottest[55,9,]
q2map[65,10,]<-q2tottest[65,10,]
q2map[75,9,]<-q2tottest[75,9,]
q2map[99,10,]<-q2tottest[99,10,]

Y2gibbs<-gibbs_sampling_map(i3grid,q2map,0.1,20,10,rv2,rh,rh,b2,3)
image(Y2gibbs[,10,])

E1=Y1gibbs-Y1
E2=Y2gibbs-Y2
E1k<-prop_kriging(E1,rh,20,20,0.1)
E2k<-prop_kriging(E2,rh,20,20,0.1)
Y1c=Y1+E1k
Y2c=Y2+E2k
facc<-Y1c
facc[]=3
facc[Y1c<q1tottest]=1
facc[Y1c>q1tottest & Y2c>q2tottest]=2
image(facc[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(grid[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(E1k[,10,],col=heat.colors(300))
image(E2k[,10,],col=heat.colors(300))
dgrid<-array(NA,c(100,20,566))
diag_lat[is.na(diag_lat)]=Inf
dgrid[5,10,]=diag_lat[,1]
dgrid[15,9,]<-diag_lat[,2]
dgrid[25,10,]<-diag_lat[,3]
dgrid[35,9,]<-diag_lat[,4]
dgrid[45,10,]<-diag_lat[,5]
dgrid[55,9,]<-diag_lat[,6]
dgrid[65,10,]<-diag_lat[,7]
dgrid[75,9,]<-diag_lat[,8]
dgrid[99,10,]<-diag_lat[,9]
Y3gibbs<-gibbs_sampling_diag(grid,dgrid,c(qd1e2,qd2e2,qd2e3),0.1,20,10,rv2,rh,rh,b2,3)
image(Y3gibbs[,10,])
facdg<-Y3gibbs
facdg[!is.na(facdg)]=3
facdg[ grid==2 & Y3gibbs<qd2e2 & Y3gibbs>qd1e2]=1
facdg[ grid==2 & Y3gibbs>qd2e2 ]=2
facdg[ grid==3 & Y3gibbs>qd2e3 ]=2
image(facdg[,10,],col=c("green","yellow","black"))
image(dgrid[,10,],col=c("green","yellow","black"))
E3=Y3gibbs-Y3
E3k<-prop_kriging(E3,rh3,20,20,0.1)
Y3c=Y3+E3k
facd<-Y3c
facd[]=3
facd[ facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=1
facd[ facc==2 & Y3c>qd2e2 ]=2
facd[ facc==3 & Y3c>qd2e3 ]=2

image(facd[,10,],col=c("green","yellow","black"))
fac_comb=facc
fac_comb[facc==2 & Y3c<qd2e2 & Y3c>qd1e2]=4
fac_comb[facc==2 & Y3c>qd2e2 ]=5
fac_comb[facc==3 & Y3c>qd2e3 ]=5
image(facc[,10,],col=c("darkslategray4","burlywood2","brown3"))
image(fac_comb[,10,],col=c("darkslategray4","burlywood2","brown3","green","yellow"))
image(dgrid[,10,],col=c("green","yellow","black"))
```

```{r}
for ( z in 1:566){
  for ( y in 1:20){
    for ( x in 1:100){
      write(toString(c(fac_comb[x,y,z])),"utahsimu5.csv",append=TRUE)
    }
  }
}
```
we also want to export the data
```{r}
grid[grid==Inf]=NA
dgrid[dgrid==Inf]=NA
comb_grid<-grid
comb_grid[dgrid==1]=4
comb_grid[dgrid==2]=5
for ( z in 1:566){
  for ( y in 1:20){
    for ( x in 1:100){
      write(toString(c(comb_grid[x,y,z])),"utahdata.csv",append=TRUE)
    }
  }
}
```
we can also compute the resulting embedded transitions on those simulations
```{r}
i1real<-i2real<-i3real<-facc
i1real[facc!=1]=0
i2real[facc==2]=1
i2real[facc!=2]=0
i3real[facc==3]=1
i3real[facc!=3]=0
r12=embedded3d(i1real,i2real)
r13=embedded3d(i1real,i3real)
r21=embedded3d(i2real,i1real)
r23=embedded3d(i2real,i3real)
r31=embedded3d(i3real,i1real)
r32=embedded3d(i3real,i2real)
matrix(c(0,r12,r13,r21,0,r23,r31,r32,0),3,3,byrow=TRUE)
r12d=embedded2d(i1,i2)
r13d=embedded2d(i1,i3)
r21d=embedded2d(i2,i1)
r23d=embedded2d(i2,i3)
r31d=embedded2d(i3,i1)
r32d=embedded2d(i3,i2)
matrix(c(0,r12d,r13d,r21d,0,r23d,r31d,r32d,0),3,3,byrow=TRUE)
```

```{r}
real3<-matrix(c(0,r12,r13,r21,0,r23,r31,r32,0),3,3,byrow=TRUE)
```
```{r}
(real1+real2+real3)/3
```